### 2023-02-24 code/processing_code/processingfile_v1.qmd
Cleaned stork data - New files in processed data

---
title: "White Stork Antimicrobial Resistance"
subtitle: "Data Cleaning"
author: "Raquel Francisco and Seth Lattner"
date: "2023-02-24"
output: html_document
---

### Processing script

This QMD file contains processing code for antimicrobial resistance data collected from White stork (*Ciconia ciconia*) in Madrid, Spain.

#### Load Packages/Data

```{r, message=FALSE, warning=FALSE}
#load required packages
library(readr)
library(dplyr)
library(tidyr)
library(skimr) 
library(here) 
```

```{r, message=FALSE, warning=FALSE}
#load in data
stork_AMR_raw <- read_csv(here::here("data","raw_data","StorkAMRraw_2020_2021.csv"))
```

Let's take a quick look at the data on hand.

```{r}
dplyr::glimpse(stork_AMR_raw)
skimr::skim(stork_AMR_raw)
```

Just based on this, it looks like there might be a bit of cleaning we'll have to do before we can get much further. Landfill use index (`mean_lui`) data are missing for all observations from the year 2020 due to the COVID-19 pandemic and will thus be excluded from future landfill use analysis. Additionally, there are data missing from the columns `ad`, `age`, `fledged`, and `hatched`.

No inputs for other variables stand out as being suspicious or blatantly wrong.

#### Data Cleaning

`NA` values in the `fledged` and `hatched` variables are more accurately represented by 0 - no eggs hatched or no eggs fledged. We will replace NA's with 0's to correct this.

```{r}
d1 <- stork_AMR_raw %>%
  tidyr::replace_na(list(fledged = 0, hatched = 0))
skimr::skim(d1)
```

We would like to add a new column for multi-drug resistance (resistance to ≥3 antibiotic classes) into our dataframe.

```{r}
#sum across antibiotic classes and create new column when sum ≥3  
d2 <- d1 %>%
  mutate(sum = rowSums(across(c(TET, SULF, AMIN, PHEN, MACR, BLACT, QUIN, POLIMYX)))) %>%
  mutate(MDR = ifelse(sum>2, 1, 0))
glimpse(d2)
```

We want to make two cleaned data files - one with only the 2021 data (which includes landfill use) and one with both 2020 and 2021 data.

```{r}
#filter out 2020 data
d3 <- d2 %>% filter(yr !=2020)
head(d3)
```

Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.

```{r}
#save 2021 data only
stork_AMR_2021 <- d3

#save for both years
stork_AMR_clean <- d2
```

#### Save Cleaned Data

```{r}
save_data_location <- here::here("data","processed_data","stork_AMR_2021.rds")
saveRDS(stork_AMR_2021, file = save_data_location)

save_data_location2 <- here::here("data","processed_data","stork_AMR_clean.rds")
saveRDS(stork_AMR_clean, file = save_data_location2)
```

### 2023-02-24 code/analysis_code/exploratory_analysis.qmd
Stork data exploration

---
title: "White Stork Antimicrobial Resistance"
subtitle: "Exploratory Analysis"
author: "Raquel Francisco and Seth Lattner"
date: "2022-02-24"
output: html_document
---

This QMD file contains exploratory analysis for antimicrobial resistance data collected from White Stork (*Ciconia ciconia*) in Madrid, Spain.

#### Load Data/Packages

```{r, warning=FALSE, message=FALSE}
#load needed packages
library(here) 
library(dplyr)
library(skimr)
library(ggplot2)
```

```{r}
#load all data
stork_AMR_clean <- readRDS(here::here("data","processed_data","stork_AMR_clean.rds"))

#load 2021 data
stork_AMR_2021 <- readRDS(here::here("data","processed_data","stork_AMR_2021.rds"))

```

#### Explore Data

First we want to make a table to take a quick look at a few variables that will be important later on.

```{r}
table1 <- stork_AMR_2021 %>%
  select(id, age, nsuccess, mean_lui, MDR) %>%
  knitr::kable()
table1
```

Plotting total MDR outcomes from both years.

```{r}
stork_AMR_clean %>%
  ggplot(aes(MDR))+
  geom_histogram()
```
Plot MDR vs age of bird

```{r}
stork_AMR_clean %>%
  ggplot(aes(as.factor(MDR), age))+
  geom_boxplot()
```
Plot MDR vs nest success

```{r}
stork_AMR_clean %>%
  ggplot(aes(as.factor(MDR), nsuccess))+
  geom_boxplot()
```
Plotting mean landfill use index vs nest success

```{r, warning=FALSE, message=FALSE}
stork_AMR_2021 %>%
  ggplot(aes(mean_lui, nsuccess))+
  geom_point()+
  geom_smooth(method = "lm")
```
Plotting mean landfill use index vs nest success, grouped by MDR outcome.

```{r, message=FALSE, warning=FALSE}
stork_AMR_2021%>%
  ggplot(aes(mean_lui, nsuccess, color=as.factor(MDR)))+
  geom_point()+
  geom_smooth(method = "lm")
```
Plotting MDR vs mean landfill use index

```{r, warning=FALSE}
stork_AMR_2021 %>%
  ggplot(aes(as.factor(MDR), mean_lui))+
  geom_boxplot()
```

There seems to be a slight decrease in the presence of MDR as landfill use index increases. Otherwise, there are not many apparent trends in the data off the bat. This exploration will help us as we move forward into a more robust statistical analysis.

### 2023-03-17 1 Data Cleaning Script/DataCleaningScript.qmd
Project Reorganization
McKay et al 2020 used as a reference

--- 2023-02-24 code/processing_code/processingfile_v1.qmd
@@
 #load in data
-stork_AMR_raw <- read_csv(here::here("data","raw_data","StorkAMRraw_2020_2021.csv"))
+stork_AMR_raw <- read_csv(here::here("1 Data Cleaning Script","raw_data","StorkAMRraw_2020_2021.csv"))
 ```
@@
 ```{r}
-save_data_location <- here::here("data","processed_data","stork_AMR_2021.rds")
+save_data_location <- here::here("2 Clean Data","stork_AMR_2021.rds")
 saveRDS(stork_AMR_2021, file = save_data_location)
 
-save_data_location2 <- here::here("data","processed_data","stork_AMR_clean.rds")
+save_data_location2 <- here::here("2 Clean Data","stork_AMR_clean.rds")
 saveRDS(stork_AMR_clean, file = save_data_location2)

### 2023-03-17 3 Analysis Scripts/Data_Exploratory_Analysis.qmd
Project Reorganization
McKay et al 2020 used as a reference

--- 2023-02-24 code/analysis_code/exploratory_analysis.qmd
@@
 #load all data
-stork_AMR_clean <- readRDS(here::here("data","processed_data","stork_AMR_clean.rds"))
+stork_AMR_clean <- readRDS(here::here("2 Clean Data","stork_AMR_clean.rds"))
 
 #load 2021 data
-stork_AMR_2021 <- readRDS(here::here("data","processed_data","stork_AMR_2021.rds"))
+stork_AMR_2021 <- readRDS(here::here("2 Clean Data","stork_AMR_2021.rds"))
 

### 2023-03-17 5 Manuscript/manuscript/Manuscript.qmd
Project Reorganization
McKay et al 2020 used as a reference

---
title: "Example Manuscript Template for a Data Analysis Project"
subtitle: ""
author: Andreas Handel
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Summary/Abstract
_Write a summary of your project._

{{< pagebreak >}}

# Introduction

## General Background Information
_Provide enough background on your topic that others can understand the why and how of your analysis_

## Description of data and data source
_We have 126 observations taken over a period of 2 years. We are evaluating for the presence and burden of antimicrobial resistance genes in white stork feces. Codebook is a WIP located in raw data folder._

## Questions/Hypotheses to be addressed
_Does landfill use by storks increase the likelihood of carrying Antimicrobial resistance (AMR) genes and resistance gene burden?._

_Stork that visit the landfill more often will have a higher likelihood of carrying AMR genes and have a higher burden of AMR genes._

_Is nest success related to AMR gene burden and multi-drug resistance (i.e., resistance to 3 or more drug classes)?_

_Stork with more AMR gene burdens and multi-drug resistance will have lower nest success._

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a]

{{< pagebreak >}}

# Methods

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._

## Data aquisition
_This study took place in Prado Herrero, a private cattle ranch located northwest of metropolitan Madrid and is surrounded by agriculture._

_Between March to May 2020 and 2021, we collected feces from marked nests with banded adults in known breeding pairs at 3 points of the breeding season; (1) an adult sample during incubation, (2) an early juvenile sample during the first two weeks of the chicks life when adults are believed to forage on natural sources, and (3) a late juvenile sample after chicks were past two weeks of age when adults forage on anthropogenic resources. Nests were visited in the late mornings and approximately one gram of fresh feces was collected from the perimeter of the nest structure into a sterile Eppendorf tube._

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._
```{r}
library(tidyverse)
library(here)

raw <- read_csv(here('data/raw_data/StorkAMRraw_2020_2021.csv'))

tibble(raw)
```

## Statistical analysis
_Basic summary statistics will be performed. Models will be built and fitted to data using the statistical package lme4 in Program R. We will evaluated model fit for all models using Akaike Information Criterion scores adjusted for small sample sizes. The best fit model was selected by having a ΔAICc < 2._

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","height_weight_stratified.png"))
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.

{{< pagebreak >}}

# References

### 2023-03-17 code/processing_code/processingcode.R

@@ -1,151 +1,0 @@
-###############################
-# processing script
-#
-#this script loads the raw data, processes and cleans it 
-#and saves it as Rds file in the processed_data folder
-#
-# Note the ## ---- name ---- notation
-# This is done so one can pull in the chunks of code into the Quarto document
-# see here: https://bookdown.org/yihui/rmarkdown-cookbook/read-chunk.html
-
-
-## ---- packages --------
-#load needed packages. make sure they are installed.
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing/cleaning
-library(tidyr) #for data processing/cleaning
-library(skimr) #for nice visualization of data 
-library(here) #to set paths
-
-## ---- loaddata --------
-#path to data
-#note the use of the here() package and not absolute paths
-data_location <- here::here("data","raw_data","exampledata.xlsx")
-
-#load data. 
-#note that for functions that come from specific packages (instead of base R)
-# I often specify both package and function like so
-#package::function() that's not required one could just call the function
-#specifying the package makes it clearer where the function "lives",
-#but it adds typing. You can do it either way.
-rawdata <- readxl::read_excel(data_location)
-
-# We can look in the Codebook (second tab) for a variable explanation
-codebook <- readxl::read_excel(data_location, sheet ="Codebook")
-print(codebook)
-
-
-## ---- exploredata --------
-#take a look at the data
-dplyr::glimpse(rawdata)
-
-#another way to look at the data
-summary(rawdata)
-
-#yet another way to get an idea of the data
-head(rawdata)
-
-#this is a nice way to look at data
-skimr::skim(rawdata)
-
-
-
-
-## ---- cleandata1 --------
-# Inspecting the data, we find some problems that need addressing:
-# First, there is an entry for height which says "sixty" instead of a number. 
-# Does that mean it should be a numeric 60? It somehow doesn't make
-# sense since the weight is 60kg, which can't happen for a 60cm person (a baby)
-# Since we don't know how to fix this, we might decide to remove the person.
-# This "sixty" entry also turned all Height entries into characters instead of numeric.
-# That conversion to character also means that our summary function isn't very meaningful.
-# So let's fix that first.
-
-d1 <- rawdata %>% dplyr::filter( Height != "sixty" ) %>% 
-                  dplyr::mutate(Height = as.numeric(Height))
-
-
-# look at partially fixed data again
-skimr::skim(d1)
-hist(d1$Height)
-
-
-## ---- cleandata2 --------
[…]

### 2023-03-18 3 Analysis Scripts/StatiscalAnalysis_ModelFitting.qmd
Model fitting and analysis
Large Issue: Data needs further cleaning. Will not add to manuscript till data is cleaned in AM.

---
title: "StatiscalAnalysis_ModelFitting"
format: html
editor: visual
---
# Ciconia ciconia analysis script

#### this script loads the processed, cleaned data, does a simple analysis and saves the results to the results folder

```{r}
#load needed packages. make sure they are installed.
library(ggplot2) #for plotting
library(ggeffects) #to explore plots  
library(broom) #for cleaning up output from lm()
library(broom.mixed) #for cleaning up output from lmer()
library(here) #for data loading/saving
library(multilevelmod) #mixed effect models
library(tidymodels)
library(lme4)
library(sjPlot) #to help plot models
library(performance) #evaluate model fit and performance

#path to data
data_location <- here::here("2 Clean Data","stork_AMR_clean.rds")

#load data. 
data <- readRDS(data_location)
```

## Data fitting/statistical analysis

Data of interest: Landfill use index (LUI) and multi-drug resistence (MDR) throughout the different sampling periods during the stork breeding season.

We'll start by fitting two models. We will used mixed-effects models to control for repeated sampling of the same nest. The design is unbalanced, thus we cannot use a repeated measures anova.

We’ll start with a single factor repeated measures evaluation the  effect of LUI on blaTEM (betalactam) resistence gene burden.
```{r}
data(data, package = "lme4")

data %>% 
  ggplot(aes(x = samp, y = blaTEM)) + 
  geom_point() + 
  geom_line() +
  facet_wrap(~ nes) 
```

Not really what we need... move on

Build models using lmer and glmer
```{r}
lmer_spec <- 
  linear_reg() %>% 
  set_engine("lmer")

glmer_spec <- 
  linear_reg() %>% 
  set_engine("glmer")

```

Multi-drug resistance (0,1 data) vs All

```{r}
lmer_fit_null <- 
  lmer_spec %>% 
  fit(MDR ~ (1|nes), data = data)

lmer_fit_global <- 
  lmer_spec %>% 
  fit(MDR ~ mean_lui + age + samp + nsuccess +(1|nes), data = data)

#Univariate models
lmer_fit_1 <- 
  lmer_spec %>% 
  fit(MDR ~ mean_lui +(1|nes), data = data)

lmer_fit_2 <- 
  lmer_spec %>% 
  fit(MDR ~ age  +(1|nes), data = data)

lmer_fit_3 <- 
  lmer_spec %>% 
  fit(MDR ~ samp +(1|nes), data = data)

lmer_fit_4 <- 
  lmer_spec %>% 
  fit(MDR ~ nsuccess +(1|nes), data = data)

lmer_fit_5 <- 
  lmer_spec %>% 
  fit(MDR ~ mean_lui + samp +(1|nes), data = data)

lmer_fit_6 <- 
  lmer_spec %>% 
  fit(MDR ~ mean_lui + nsuccess +(1|nes), data = data)

lmer_fit_7 <- 
  lmer_spec %>% 
  fit(MDR ~ samp + nsuccess +(1|nes), data = data)

lmer_fit_8 <- 
  lmer_spec %>% 
  fit(MDR ~ mean_lui*age + samp + nsuccess +(1|nes), data = data)

lmer_fit_9 <- 
  lmer_spec %>% 
  fit(MDR ~ mean_lui + age*samp + nsuccess +(1|nes), data = data)

lmer_fit_10 <- 
  lmer_spec %>% 
  fit(MDR ~ mean_lui + age + samp*nsuccess +(1|nes), data = data)

```

Use AIC to see the best model; None look great...

```{r}
AIC_table1 <- compare_performance(lmer_fit_null,lmer_fit_global, lmer_fit_1, lmer_fit_2, lmer_fit_3, lmer_fit_4, lmer_fit_5, lmer_fit_6, lmer_fit_7, lmer_fit_8, lmer_fit_9, lmer_fit_10)

print(AIC_table1)
AIC_table1 = here("4 Results","Tables", "ModelAIC1.rds")
saveRDS(AIC_table1, file = AIC_table1)

plot(compare_performance(lmer_fit_null,lmer_fit_global, lmer_fit_1, lmer_fit_2, lmer_fit_3, lmer_fit_4, lmer_fit_5, lmer_fit_6, lmer_fit_7, lmer_fit_8, lmer_fit_9, lmer_fit_10, rank = TRUE))
```
Summarize and graph best model (lowwest AIC)

```{r}
lmer_fit_global

# place results from fit into a data frame with the tidy function

model1 <- broom::tidy(lmer_fit_global)
print(model1)

# save fit results table  
model1 = here("4 Results", "Models", "LMERmodel1.rds")
saveRDS(model1, file = model1)

MLp <- plot_model(lmer_fit_global,terms = 'mean_lui', type="pred", 
           show.data = TRUE, show.values = TRUE, show.p = TRUE,
           title = "", show.legend = TRUE, line.size=1,
           axis.title = c("Landfill Use Index",'Multi-Drug Resistence'))

MLp
```

Class Diversity (sum) vs All

```{r}
lmer_fit2_null <- #singular
  lmer_spec %>% 
  fit(sum ~ (1|nes), data = data)

lmer_fit2_global <- 
  lmer_spec %>% 
  fit(sum ~ mean_lui + age + samp + nsuccess +(1|nes), data = data)

#Univariate models
lmer_fit2_1 <- 
  lmer_spec %>% 
  fit(sum ~ mean_lui +(1|nes), data = data)

lmer_fit2_2 <- #singular
  lmer_spec %>% 
  fit(sum ~ age  +(1|nes), data = data)

lmer_fit2_3 <- #singular
  lmer_spec %>% 
  fit(sum ~ samp +(1|nes), data = data)

lmer_fit2_4 <- #singular
  lmer_spec %>% 
  fit(sum ~ nsuccess +(1|nes), data = data)

lmer_fit2_5 <- 
  lmer_spec %>% 
  fit(sum ~ mean_lui + samp +(1|nes), data = data)

lmer_fit2_6 <- 
  lmer_spec %>% 
  fit(sum ~ mean_lui + nsuccess +(1|nes), data = data)

lmer_fit2_7 <- #singular
  lmer_spec %>% 
  fit(sum ~ samp + nsuccess +(1|nes), data = data)

lmer_fit2_8 <- 
  lmer_spec %>% 
  fit(sum ~ mean_lui*age + samp + nsuccess +(1|nes), data = data)

lmer_fit2_9 <- 
  lmer_spec %>% 
  fit(sum ~ mean_lui + age*samp + nsuccess +(1|nes), data = data)

lmer_fit2_10 <- 
  lmer_spec %>% 
  fit(sum ~ mean_lui + age + samp*nsuccess +(1|nes), data = data)

```

Use AIC to see the best model; This doesnt look great either...

```{r}
AIC_table2 <- compare_performance(lmer_fit2_null,lmer_fit2_global, lmer_fit2_1, lmer_fit2_2, lmer_fit2_3, lmer_fit2_4, lmer_fit2_5, lmer_fit2_6, lmer_fit2_7, lmer_fit2_8, lmer_fit2_9, lmer_fit2_10)

print(AIC_table2)
AIC_table2 = here("4 Results","Tables", "ModelAIC2.rds")
saveRDS(AIC_table2, file = AIC_table2)

plot(compare_performance(lmer_fit2_null,lmer_fit2_global, lmer_fit2_1, lmer_fit2_2, lmer_fit2_3, lmer_fit2_4, lmer_fit2_5, lmer_fit2_6, lmer_fit2_7, lmer_fit2_8, lmer_fit2_9, lmer_fit2_10, rank = TRUE))
```
Summarize and graph best model (lowwest AIC)

```{r}
lmer_fit2_10
# place results from fit into a data frame with the tidy function

model2 <- broom::tidy(lmer_fit2_10)
print(model2)

# save fit results table  
model2 = here("4 Results","Models", "LMERmodel2.rds")
saveRDS(model2, file = model2)

CLp <- plot_model(lmer_fit2_10,terms = 'mean_lui', type="pred", 
           show.data = TRUE, show.values = TRUE, show.p = TRUE,
           title = "", show.legend = TRUE, line.size=1,
           axis.title = c("Landfill Use Index",'Total Antibiotic Classes with Resistance'))

CLp
```

## Pause... Big issue. Models don't appear to be working well... Have to go back and reclean the data. Covariates should be standarized; values too large between variables! Correlation between variables should be analyzed, plan to use spearman. Will complete 3/17/23

### 2023-03-18 3 Analysis Scripts/Statistical_Analysis.R

@@ -1,56 +1,0 @@
-###############################
-# analysis script
-#
-#this script loads the processed, cleaned data, does a simple analysis
-#and saves the results to the results folder
-
-#load needed packages. make sure they are installed.
-library(ggplot2) #for plotting
-library(broom) #for cleaning up output from lm()
-library(here) #for data loading/saving
-
-#path to data
-#note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed_data","processeddata.rds")
-
-#load data. 
-mydata <- readRDS(data_location)
-
-
-######################################
-#Data fitting/statistical analysis
-######################################
-
-############################
-#### First model fit
-# fit linear model using height as outcome, weight as predictor
-
-lmfit1 <- lm(Height ~ Weight, mydata)  
-
-# place results from fit into a data frame with the tidy function
-lmtable1 <- broom::tidy(lmfit1)
-
-#look at fit results
-print(lmtable1)
-
-# save fit results table  
-table_file1 = here("results", "resulttable1.rds")
-saveRDS(lmtable1, file = table_file1)
-
-############################
-#### Second model fit
-# fit linear model using height as outcome, weight and gender as predictor
-
-lmfit2 <- lm(Height ~ Weight + Gender, mydata)  
-
-# place results from fit into a data frame with the tidy function
-lmtable2 <- broom::tidy(lmfit2)
-
-#look at fit results
-print(lmtable2)
-
-# save fit results table  
-table_file2 = here("results", "resulttable2.rds")
-saveRDS(lmtable2, file = table_file2)
-
-  

### 2023-03-20 1 Data Cleaning Script/DataCleaningScript.qmd
Pt.2 Model Fitting
Data re-cleaned, re-explored, and fitted to models

@@
 
+Standardize main continuous covariates (mean_lui, nsuccess, age)
+
+```{r}
+d2$s.nsuccess <- (d2$nsuccess-mean(d2$nsuccess, na.rm = TRUE))/sd(d2$nsuccess, na.rm = TRUE)
+
+#?sd, use na.rm = TRUE to ensure that NAs are removed from calculations
+d2$s.lui <- (d2$mean_lui-mean(d2$mean_lui, na.rm = TRUE))/sd(d2$mean_lui, na.rm = TRUE)
+
+d2$s.age <- (d2$age-mean(d2$age, na.rm = TRUE))/sd(d2$age, na.rm = TRUE)
+
+```
+
+Quantitative data is reported in log10.
+Will transform columns to data with the inverse log10 that would allow a more accurate evaluation of AMR gene burden via formula; x=10^y where x is the new gene burden and y is the original gene burden calculated from published formula.
+
+```{r}
+d2$TET..A <- 10^(d2$TET..A)
+d2$TET..B <- 10^(d2$TET..B)
+d2$TET..Y <- 10^(d2$TET..Y)
+d2$TET..K <- 10^(d2$TET..K)
+d2$TET..M <- 10^(d2$TET..M)
+d2$TET..Q <- 10^(d2$TET..Q)
+d2$TET..S <- 10^(d2$TET..S)
+d2$TET..W <- 10^(d2$TET..W)
+d2$sulI <- 10^(d2$sulI)
+d2$sulII <- 10^(d2$sulII)
+d2$str <- 10^(d2$str)
+d2$aadA <- 10^(d2$aadA)
+d2$catI <- 10^(d2$catI)
+d2$catII <- 10^(d2$catII)
+d2$ermB <- 10^(d2$ermB)
+d2$ermF <- 10^(d2$ermF)
+d2$qnrB <- 10^(d2$qnrB)
+d2$qnrS <- 10^(d2$qnrS)
+d2$blaTEM <- 10^(d2$blaTEM)
+d2$mecA <- 10^(d2$mecA)
+d2$mcr.1 <- 10^(d2$mcr.1)
+
+d3 <- d2 %>%
+  mutate(Burden = rowSums(across(c(TET..A, TET..B, TET..Y, TET..K, TET..M, TET..Q, TET..S, TET..W, sulI, sulII, str, aadA, catI, catII, ermB, ermF, qnrB, qnrS, blaTEM, mecA, mcr.1))))
+
+glimpse(d3)
+```
+
 We want to make two cleaned data files - one with only the 2021 data (which includes landfill use) and one with both 2020 and 2021 data.
@@
 #filter out 2020 data
-d3 <- d2 %>% filter(yr !=2020)
-head(d3)
+d4 <- d3 %>% filter(yr !=2020)
+head(d4)
 ```
@@
 #save 2021 data only
-stork_AMR_2021 <- d3
+stork_AMR_2021 <- d4
 
 #save for both years
-stork_AMR_clean <- d2
+stork_AMR_clean <- d3
 ```

### 2023-03-20 3 Analysis Scripts/Data_Exploratory_Analysis.qmd
Pt.2 Model Fitting
Data re-cleaned, re-explored, and fitted to models

@@
 #load needed packages
-library(here) 
-library(dplyr)
+library(here) #calling
+library(dplyr) #cleaning
 library(skimr)
-library(ggplot2)
+library(ggplot2) #plotting and visualization
+library(ggpubr) #ggscatter to visualize correlations
+library(naniar)#missing values graphics and plots
+library(UpSetR)#missing values graphics and plots
+library(corrplot) #evaluate if covariates are correlated
 ```
@@
 
-#### Explore Data
+#### Quality Control
+
+```{r}
+stork_AMR_clean %>%
+  as_shadow_upset() %>%
+  upset()
+
+gg_miss_fct(x = stork_AMR_clean, fct = yr)
+gg_miss_fct(x = stork_AMR_clean, fct = samp)
+```
+Quite alot of missing values for 2020. Will likely only be able to do summary stats on 2020 because no LUI data was collected that year (COVID-19 limitations)
+
+#### Data subset for 2021
+
+```{r}
+stork_AMR_2021 %>%
+  as_shadow_upset() %>%
+  upset()
+
+gg_miss_fct(x = stork_AMR_2021, fct = yr)
+gg_miss_fct(x = stork_AMR_2021, fct = samp)
+```
+
+#### Correlation Check of Covariates
+
+```{r}
+#define the variables we want to include in the correlation matrix
+variables <- stork_AMR_clean[ , c("s.age", "s.lui", "samp", "s.nsuccess")]
+
+#create correlation matrix
+cor(variables, use = "complete.obs") #ignore NAs
+corrplot(cor(variables, use = "complete.obs"))
+
+Z <- data.frame(s.age = stork_AMR_clean$s.age, 
+                s.lui = stork_AMR_clean$s.lui, 
+                s.nsuccess = stork_AMR_clean$s.nsuccess,
+                samp = stork_AMR_clean$samp)
+pairs(Z, panel = panel.smooth, labels = c("age", "mean_lui", "nsuccess", "samp"))
+
+#More than 0.7 usually is unacceptable
+#More than 0.5 you need to be careful; NOTHING over 0.5
+
+##BASIC SPEARMAN
+cor.test( stork_AMR_clean$s.age, stork_AMR_clean$s.lui,
+         method = "spearman")
+cor.test( stork_AMR_clean$s.age, stork_AMR_clean$s.nsuccess,
+          method = "spearman")
+cor.test( stork_AMR_clean$s.age, stork_AMR_clean$samp,
+          method = "spearman")
+cor.test( stork_AMR_clean$s.nsuccess, stork_AMR_clean$s.lui,
+          method = "spearman")
+cor.test( stork_AMR_clean$s.nsuccess, stork_AMR_clean$samp,
+          method = "spearman")
+cor.test( stork_AMR_clean$s.lui, stork_AMR_clean$samp,
+          method = "spearman")
+cor.test( stork_AMR_clean$Burden, stork_AMR_clean$MDR,
+          method = "spearman")
+cor.test( stork_AMR_clean$sum, stork_AMR_clean$MDR,
+          method = "spearman") #Correlated!
+
+##Visual Plots with: regression line and confidence interval
+
+#Age + LUI
+ggscatter(stork_AMR_clean, x = 'age', y = 's.lui',
+          add = "reg.line", conf.int = TRUE,    
+          add.params = list(fill = "lightgray"),
+          ggtheme = theme_minimal()
+)+
+  stat_cor(method = "spearman", p.accuracy = 0.001, r.accuracy = 0.01,
+           label.y = 0.5)
+
+#Nsuccess + LUI
+ggscatter(stork_AMR_clean, x = 's.nsuccess', y = 's.lui',
+          add = "reg.line", conf.int = TRUE,    
+          add.params = list(fill = "lightgray"),
+          ggtheme = theme_minimal()
+)+
+  stat_cor(method = "spearman", p.accuracy = 0.001, r.accuracy = 0.01,
+           label.y = 0.6)
+
+#Nsuccess + Age
+ggscatter(stork_AMR_clean, x = 's.nsuccess', y = 'age',
+          add = "reg.line", conf.int = TRUE,    
+          add.params = list(fill = "lightgray"),
+          ggtheme = theme_minimal()
+)+
+  stat_cor(method = "spearman", p.accuracy = 0.001, r.accuracy = 0.01,
+           label.y = 20)
+
+#Variables of interest not correlated; they are independent = Move on to analysis
+```
+
+#### Data Exploration and visualization
+
+Will explore both 2020 and 2021 data, but will not model with 2020 as no LUI data was collected this year due to COVID-19 restrictions.
 
@@
 ```{r}
-table1 <- stork_AMR_2021 %>%
-  select(id, age, nsuccess, mean_lui, MDR) %>%
+table1 <- stork_AMR_clean %>%
+  select(id, age, nsuccess, mean_lui, Burden, MDR) %>%
   knitr::kable()
@@
 
-Plotting total MDR outcomes from both years.
+Plotting total Burden outcomes from both years.
 
@@
 stork_AMR_clean %>%
-  ggplot(aes(MDR))+
-  geom_histogram()
+  ggplot(aes(as.factor(yr), Burden))+
+  geom_boxplot()
 ```
+Plotting total Burden outcomes across all sampling periods.
+```{r}
+stork_AMR_clean %>%
+  ggplot(aes(as.factor(samp), Burden))+
+  geom_boxplot()
+```
+Total Number of AMR Gene Classes over both years.
+
+```{r}
+stork_AMR_clean %>%
+  ggplot(aes(as.factor(yr), sum))+
+  geom_boxplot()
+```
+
+Sample vs Total Number of AMR Gene Classes
+```{r}
+stork_AMR_clean %>%
+  ggplot(aes(as.factor(samp), sum))+
+  geom_boxplot()
+```
+
+Plot Gene Burden vs age of bird
+
+```{r}
+stork_AMR_clean %>%
+  ggplot(aes(age, Burden))+
+  geom_point()
+```
+
 Plot MDR vs age of bird
@@
 ```
+
+Plot Gene Burden vs age of bird
+
+```{r}
+stork_AMR_clean %>%
+  ggplot(aes(nsuccess, Burden))+
+  geom_point()
+```
+
 Plot MDR vs nest success
@@
 ```
+
+Now we will make a table filtering out 2020.
+
+```{r}
+table2 <- stork_AMR_2021 %>%
+  select(id, age, nsuccess, mean_lui, Burden, MDR) %>%
+  knitr::kable()
+table2
+```
+
 Plotting mean landfill use index vs nest success
@@
 ```
+
 Plotting mean landfill use index vs nest success, grouped by MDR outcome.
@@
 ```
+
+Plotting mean landfill use index vs nest success, grouped by MDR outcome.
+
+```{r, message=FALSE, warning=FALSE}
+stork_AMR_2021%>%
+  ggplot(aes(mean_lui, nsuccess, color=Burden))+
+  geom_point()+
+  geom_smooth(method = "lm")
+```
+
 Plotting MDR vs mean landfill use index
@@
 
-There seems to be a slight decrease in the presence of MDR as landfill use index increases.
+Plotting Burden vs mean landfill use index
+
+```{r, warning=FALSE}
+stork_AMR_2021 %>%
+  ggplot(aes(mean_lui, Burden))+
+  geom_point()+
+  geom_smooth(method = "lm")
+```
+
+There seems to be a slight decrease in the presence of MDR and increase in the burden of AMR genes as landfill use index increases.
 Otherwise, there are not many apparent trends in the data off the bat.

### 2023-03-20 3 Analysis Scripts/StatiscalAnalysis_ModelFitting.qmd
Pt.2 Model Fitting
Data re-cleaned, re-explored, and fitted to models

@@
 ---
+
 # Ciconia ciconia analysis script
@@
 #path to data
-data_location <- here::here("2 Clean Data","stork_AMR_clean.rds")
+data_location <- here::here("2 Clean Data","stork_AMR_2021.rds")
 
@@
 
-We’ll start with a single factor repeated measures evaluation the  effect of LUI on blaTEM (betalactam) resistence gene burden.
+We'll start with a single factor repeated measures evaluation the effect of LUI on blaTEM (betalactam) resistence gene burden.
+
 ```{r}
@@
 data %>% 
-  ggplot(aes(x = samp, y = blaTEM)) + 
+  ggplot(aes(x = samp, y = Burden)) + 
   geom_point() + 
@@
 
-Not really what we need... move on
+No obvious trends...
 
 Build models using lmer and glmer
+
 ```{r}
@@
   lmer_spec %>% 
-  fit(MDR ~ mean_lui + age + samp + nsuccess +(1|nes), data = data)
+  fit(MDR ~ s.lui + s.age + samp + s.nsuccess +(1|nes), data = data)
 
@@
   lmer_spec %>% 
-  fit(MDR ~ mean_lui +(1|nes), data = data)
+  fit(MDR ~ s.lui +(1|nes), data = data)
 
@@
   lmer_spec %>% 
-  fit(MDR ~ age  +(1|nes), data = data)
+  fit(MDR ~ s.age  +(1|nes), data = data)
 
@@
   lmer_spec %>% 
-  fit(MDR ~ nsuccess +(1|nes), data = data)
+  fit(MDR ~ s.nsuccess +(1|nes), data = data)
 
@@
   lmer_spec %>% 
-  fit(MDR ~ mean_lui + samp +(1|nes), data = data)
+  fit(MDR ~ s.lui + samp +(1|nes), data = data)
 
@@
   lmer_spec %>% 
-  fit(MDR ~ mean_lui + nsuccess +(1|nes), data = data)
+  fit(MDR ~ s.lui + s.nsuccess +(1|nes), data = data)
 
@@
   lmer_spec %>% 
-  fit(MDR ~ samp + nsuccess +(1|nes), data = data)
+  fit(MDR ~ samp + s.nsuccess +(1|nes), data = data)
 
@@
   lmer_spec %>% 
-  fit(MDR ~ mean_lui*age + samp + nsuccess +(1|nes), data = data)
+  fit(MDR ~ s.lui*s.age + samp + s.nsuccess +(1|nes), data = data)
 
@@
   lmer_spec %>% 
-  fit(MDR ~ mean_lui + age*samp + nsuccess +(1|nes), data = data)
+  fit(MDR ~ s.lui + s.age*samp + s.nsuccess +(1|nes), data = data)
 
@@
   lmer_spec %>% 
-  fit(MDR ~ mean_lui + age + samp*nsuccess +(1|nes), data = data)
+  fit(MDR ~ s.lui + s.age + samp*s.nsuccess +(1|nes), data = data)
 
+lmer_fit_11 <- 
+  lmer_spec %>% 
+  fit(MDR ~ s.lui + s.age +(1|nes), data = data)
+
+lmer_fit_12 <- 
+  lmer_spec %>% 
+  fit(MDR ~ s.lui + s.age +  s.nsuccess +(1|nes), data = data)
 ```
@@
 ```{r}
-AIC_table1 <- compare_performance(lmer_fit_null,lmer_fit_global, lmer_fit_1, lmer_fit_2, lmer_fit_3, lmer_fit_4, lmer_fit_5, lmer_fit_6, lmer_fit_7, lmer_fit_8, lmer_fit_9, lmer_fit_10)
+AIC_table1 <- compare_performance(lmer_fit_null,lmer_fit_global, lmer_fit_1, lmer_fit_2, lmer_fit_3, lmer_fit_4, lmer_fit_5, lmer_fit_6, lmer_fit_7, lmer_fit_8, lmer_fit_9, lmer_fit_10, lmer_fit_11, lmer_fit_12)
 
@@
 
-plot(compare_performance(lmer_fit_null,lmer_fit_global, lmer_fit_1, lmer_fit_2, lmer_fit_3, lmer_fit_4, lmer_fit_5, lmer_fit_6, lmer_fit_7, lmer_fit_8, lmer_fit_9, lmer_fit_10, rank = TRUE))
+plot(compare_performance(lmer_fit_null,lmer_fit_global, lmer_fit_1, lmer_fit_2, lmer_fit_3, lmer_fit_4, lmer_fit_5, lmer_fit_6, lmer_fit_7, lmer_fit_8, lmer_fit_9, lmer_fit_10, lmer_fit_11, lmer_fit_12, rank = TRUE))
 ```
+
 Summarize and graph best model (lowwest AIC)
@@
 ```{r}
-lmer_fit_global
+check_model(lmer_fit_11)
 
@@
 
-model1 <- broom::tidy(lmer_fit_global)
+model1 <- broom::tidy(lmer_fit_11)
 print(model1)
@@
 
-MLp <- plot_model(lmer_fit_global,terms = 'mean_lui', type="pred", 
+MLp <- plot_model(lmer_fit_11,terms = 's.lui', type="pred", 
            show.data = TRUE, show.values = TRUE, show.p = TRUE,
@@
 MLp
+
+ggsave(MLp, file="LUIvsMDR_LME.tiff")
 ```
 
-Class Diversity (sum) vs All
+AMR Gene Burden vs All
 
@@
   lmer_spec %>% 
-  fit(sum ~ (1|nes), data = data)
+  fit(Burden ~ (1|nes), data = data)
 
@@
   lmer_spec %>% 
-  fit(sum ~ mean_lui + age + samp + nsuccess +(1|nes), data = data)
+  fit(Burden ~ s.lui + s.age + samp + s.nsuccess +(1|nes), data = data) #overfit
 
@@
   lmer_spec %>% 
-  fit(sum ~ mean_lui +(1|nes), data = data)
+  fit(Burden ~ s.lui +(1|nes), data = data)
 
-lmer_fit2_2 <- #singular
+lmer_fit2_2 <- 
   lmer_spec %>% 
-  fit(sum ~ age  +(1|nes), data = data)
+  fit(Burden ~ s.age  +(1|nes), data = data) #overfit
 
-lmer_fit2_3 <- #singular
+lmer_fit2_3 <- 
   lmer_spec %>% 
-  fit(sum ~ samp +(1|nes), data = data)
+  fit(Burden ~ samp +(1|nes), data = data)
 
-lmer_fit2_4 <- #singular
+lmer_fit2_4 <- 
   lmer_spec %>% 
-  fit(sum ~ nsuccess +(1|nes), data = data)
+  fit(Burden ~ s.nsuccess +(1|nes), data = data)
 
@@
   lmer_spec %>% 
-  fit(sum ~ mean_lui + samp +(1|nes), data = data)
+  fit(Burden ~ s.lui + samp +(1|nes), data = data)
 
@@
   lmer_spec %>% 
-  fit(sum ~ mean_lui + nsuccess +(1|nes), data = data)
+  fit(Burden ~ s.lui + s.nsuccess +(1|nes), data = data)
 
-lmer_fit2_7 <- #singular
+lmer_fit2_7 <- 
   lmer_spec %>% 
-  fit(sum ~ samp + nsuccess +(1|nes), data = data)
+  fit(Burden ~ samp + s.nsuccess +(1|nes), data = data)
 
@@
   lmer_spec %>% 
-  fit(sum ~ mean_lui*age + samp + nsuccess +(1|nes), data = data)
+  fit(Burden ~ s.lui*s.age + samp + s.nsuccess +(1|nes), data = data) #overfit
 
@@
   lmer_spec %>% 
-  fit(sum ~ mean_lui + age*samp + nsuccess +(1|nes), data = data)
+  fit(Burden ~ s.lui + s.age*samp + s.nsuccess +(1|nes), data = data) #overfit
 
@@
   lmer_spec %>% 
-  fit(sum ~ mean_lui + age + samp*nsuccess +(1|nes), data = data)
+  fit(Burden ~ s.lui + s.age + samp*s.nsuccess +(1|nes), data = data) #overfit
 
+lmer_fit2_11 <- 
+  lmer_spec %>% 
+  fit(Burden ~ s.lui + s.age + s.nsuccess +(1|nes), data = data) #overfit
+
+lmer_fit2_12 <- 
+  lmer_spec %>% 
+  fit(Burden ~ s.lui + s.age +(1|nes), data = data) #overfit
+
+#Standardizing variables helped with the overfit/singular model results
 ```
@@
 ```{r}
-AIC_table2 <- compare_performance(lmer_fit2_null,lmer_fit2_global, lmer_fit2_1, lmer_fit2_2, lmer_fit2_3, lmer_fit2_4, lmer_fit2_5, lmer_fit2_6, lmer_fit2_7, lmer_fit2_8, lmer_fit2_9, lmer_fit2_10)
+AIC_table2 <- compare_performance(lmer_fit2_null,lmer_fit2_global, lmer_fit2_1, lmer_fit2_2, lmer_fit2_3, lmer_fit2_4, lmer_fit2_5, lmer_fit2_6, lmer_fit2_7, lmer_fit2_8, lmer_fit2_9, lmer_fit2_10, lmer_fit2_11, lmer_fit2_12)
 
@@
 
-plot(compare_performance(lmer_fit2_null,lmer_fit2_global, lmer_fit2_1, lmer_fit2_2, lmer_fit2_3, lmer_fit2_4, lmer_fit2_5, lmer_fit2_6, lmer_fit2_7, lmer_fit2_8, lmer_fit2_9, lmer_fit2_10, rank = TRUE))
+plot(compare_performance(lmer_fit2_null,lmer_fit2_global, lmer_fit2_1, lmer_fit2_2, lmer_fit2_3, lmer_fit2_4, lmer_fit2_5, lmer_fit2_6, lmer_fit2_7, lmer_fit2_8, lmer_fit2_9, lmer_fit2_10, lmer_fit2_11, lmer_fit2_12, rank = TRUE))
 ```
+
 Summarize and graph best model (lowwest AIC)
@@
 ```{r}
-lmer_fit2_10
-# place results from fit into a data frame with the tidy function
+check_model(lmer_fit2_5)
 
-model2 <- broom::tidy(lmer_fit2_10)
+model2 <- broom::tidy(lmer_fit2_5)
 print(model2)
@@
 
-CLp <- plot_model(lmer_fit2_10,terms = 'mean_lui', type="pred", 
+BLp <- plot_model(lmer_fit2_5,terms = 's.lui', type="pred", 
            show.data = TRUE, show.values = TRUE, show.p = TRUE,
            title = "", show.legend = TRUE, line.size=1,
-           axis.title = c("Landfill Use Index",'Total Antibiotic Classes with Resistance'))
+           axis.title = c("Landfill Use Index",'Total Antibiotic Resistant Gene Burden'))
 
-CLp
+BLp
+
+ggsave(BLp, file="LUIvsBurden_LME.tiff")
 ```
 
-## Pause...
-Big issue.
-Models don't appear to be working well...
-Have to go back and reclean the data.
-Covariates should be standarized; values too large between variables!
-Correlation between variables should be analyzed, plan to use spearman.
-Will complete 3/17/23
+Preliminary results
+
+Questions: 
+Should over fit models be taken out of performance analysis?
+What is the best tool/package to control for small data sets?

### 2023-03-20 5 Manuscript/manuscript/Manuscript.qmd
Pt.2 Model Fitting
Data re-cleaned, re-explored, and fitted to models

@@
 ---
-title: "Example Manuscript Template for a Data Analysis Project"
-subtitle: ""
-author: Andreas Handel
+title: "A Match Made in Landfills?"
+subtitle: "Exploring the diversity and burden of antimicrobial resistance genes carried by white stork (Ciconia ciconia) throughout the breeding season in Madrid, Spain"
+author: Raquel Francisco and Seth Latter
 date: "`r Sys.Date()`"
@@
 
-The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
-You don't need to have exactly these sections, but the content covering those sections should be addressed.
-
-This uses MS Word as output format.
-[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
-You can switch to other formats, like html or pdf.
-See [the Quarto documentation](https://quarto.org/) for other formats.
-
 ```{r, echo=FALSE, message=FALSE}
@@
 
-# Summary/Abstract
-_Write a summary of your project._
+# Highlights
+
+- Our results suggest that landfills may not be as impactful as previously believed to the emergence and maintenance of AMR in this system.
+
+- Storks are most impacted by multi-drug resistance and ARG burden when anthropogenic waste is most heavily consumed later in the breeding season.
+
+- This species appears to be a good sentinel for anthropogenic impact on environment.
+
+- Future efforts in stork AMR research should focus on exploring the relationship between other anthropogenic environments (e.g., agricultural pastures) and health.
+
+# Abstract
+
+Anthropogenic environments are hotspots for the emergence and maintenance of antimicrobial resistance.
+Agricultural pastures and landfills are of particular interest due to their complex microbial communities and abundant wildlife visitation, which could facilitate the exchange of antimicrobial resistance genes (ARGs) via horizontal transfer.
+Wild birds that occupy these environments may become both reservoirs and transporters of ARGs.
+The White Stork is a highly urbanized wading bird that has significantly changed its ecology due to shifts in Spanish waste management toward open-air landfills.
+This species now heavily forages at landfills, which provide abundant food and allow for improved reproductive success.
+Birds that are dependent on anthropogenic resources, such as storks, provide the ideal opportunity to understand the emergence and spread of AMR.
+We evaluated the diversity and quantity of ARGs in storks during three periods of the breeding season (defined by distinct foraging strategies).
+A total of 31 nests at Prado Herrero were sampled between March-July (2020-2021).
+Fresh feces were collected from 31 nests to evaluate the presence of 23 important ARGs affecting eight antibiotic classes via quantitative PCR.
+All nests carried multiple ARGs.
+Over 70% of nests had multi-drug resistance to at least 3 antibiotic classes during at least one time-period.
+Generalized Linear Mixed Models revealed that increased diversity in antibiotic class resistance, amount of ARGs present in a sample, and multi-drug resistance were associated with increased adult age and decreased landfill use.
+Our results suggest that landfills may not be contributing significantly to the emergence and maintenance of AMR in this system.
+Little literature exists on the relationship between stork habitat selection and health outside of landfill use in Spain.
+Future efforts in stork AMR research should focus on exploring the relationship between agricultural land use and health.
 
@@
 ## General Background Information
-_Provide enough background on your topic that others can understand the why and how of your analysis_
+
+*Provide enough background on your topic that others can understand the why and how of your analysis*
 
 ## Description of data and data source
-_We have 126 observations taken over a period of 2 years.
+
+*We have 126 observations taken over a period of 2 years.
 We are evaluating for the presence and burden of antimicrobial resistance genes in white stork feces.
-Codebook is a WIP located in raw data folder._
+Codebook is a WIP located in raw data folder.*
 
 ## Questions/Hypotheses to be addressed
-_Does landfill use by storks increase the likelihood of carrying Antimicrobial resistance (AMR) genes and resistance gene burden?._
 
-_Stork that visit the landfill more often will have a higher likelihood of carrying AMR genes and have a higher burden of AMR genes._
+Does landfill use by storks increase the likelihood of carrying Antimicrobial resistance (AMR) genes and resistance gene burden?.
 
-_Is nest success related to AMR gene burden and multi-drug resistance (i.e., resistance to 3 or more drug classes)?_
+Stork that visit the landfill more often will have a higher likelihood of carrying AMR genes and have a higher burden of AMR genes.
 
-_Stork with more AMR gene burdens and multi-drug resistance will have lower nest success._
+Is nest success related to AMR gene burden and multi-drug resistance (i.e., resistance to 3 or more drug classes)?
 
-To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key.
-Then you can include like this:
+Stork with more AMR gene burdens and multi-drug resistance will have lower nest success.
+
+*To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key.
+Then you can include like this:*
 
@@
 
-_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._
+## Data aquisition
 
-## Data aquisition
-_This study took place in Prado Herrero, a private cattle ranch located northwest of metropolitan Madrid and is surrounded by agriculture._
+### Experimental Design
 
-_Between March to May 2020 and 2021, we collected feces from marked nests with banded adults in known breeding pairs at 3 points of the breeding season; (1) an adult sample during incubation, (2) an early juvenile sample during the first two weeks of the chicks life when adults are believed to forage on natural sources, and (3) a late juvenile sample after chicks were past two weeks of age when adults forage on anthropogenic resources.
-Nests were visited in the late mornings and approximately one gram of fresh feces was collected from the perimeter of the nest structure into a sterile Eppendorf tube._
+*Study Area* — This study took place in Prado Herrero, a private cattle ranch located northwest of metropolitan Madrid and is surrounded by agriculture (e.g.
+beef cattle, cereal grains, legumes, and forage plants (“Renta Agraria,” 2019)).
+Prado Herrero is located within a nationally protected area (Cuenca Alta del Manzanares Regional Park) and is just north of Santillana Reservoir.
+This reservoir that was declared an Important Bird Area by Regional Catalogue of Reservoirs and Wetlands of the Community of Madrid due to the numerous resident and migratory species that utilize this water source <CITE>.
+This cattle ranch has supported a productive white stork rookery where storks have been banded and monitored by biologists at UCM for over 20 years <CITE: Aguirre and Atienza>.
+During the 2020 to 2021 breeding season, between the months of March to June, stork nests were identified, marked, and monitored for productivity.
+Of marked nests between the 2020 and 2021 breeding seasons, 31 with banded adults were used both years to lay eggs successfully.
+All 31 nests were located in ash trees (Fraxinus angustifolius ) found within the cattle pasture.
+Storks that breed within the Prado Herrero rookery are known to utilize Colmenar Viejo Landfill which is located approximately 12km southeast.
+Colmenar Viejo is an open-air landfill and it is second largest of it’s kind in the Madrid region <CITE: Alejandro 2022>.
+
+*Sample Collection* — Between March to May 2020 and 2021, we collected feces from marked nests with banded adults in known breeding pairs at 3 points of the breeding season; (1) an adult sample during incubation, (2) an early juvenile sample during the first two weeks of the chicks life when adults are believed to forage on natural sources  <CITE: >, and (3) a late juvenile sample after chicks were past two weeks of age when adults forage on anthropogenic resources.
+Nests were visited in the late mornings and approximately one gram of fresh feces was collected from the perimeter of the nest structure into a sterile Eppendorf tube ().
+Samples were maintained cold in a portable cooler with frozen gel packs and frozen in a −20°C freezer within 4 hours of collection and processed at a later date.
+
+*Ethics statement*: All animal handling was authorized by Cumunidad de Madrid: Consejeria de Medio Ambiente, Administracion Local y Ordenacion de Territorio.
+The permit number is  D.N.I.
+nº 07.239.972-D.
+
+### Molecular analysis of ARGs
+We performed total DNA extraction directly from fecal samples, by using a pressure filtration technique (QuickGene DNA Tissue Kit S, Fujifilm, Japan) following the manufacturer’s instructions.
+The 16S rRNA gene was amplified in each DNA sample by real time PCR (rtPCR) in 10-fold dilutions of extracted samples, according to Jiang et al.
+(2013)<CITE:Fix>.
+A DNA sample was considered validated when a ten-fold dilution showed a cycle threshold (Ct) less than 25 (Esperón et al., 2018).
+Once validated, we analyzed DNA samples by with a panel of 21 different ARGs encoding resistance to eight different antimicrobial classes: tetracyclines (tet(A), tet(B), tet(Y), tet(K), tet(M), tet(Q), tet(S), and tet(W)), sulfonamides (sulI and sulII), aminoglycosides (str and aadA), phenicols (catI and catII), macrolides (ermB and ermF), quinolones (qnrS and qnrB), betalactams (blaTEM and mecA), and polymyxins (mcr-1).
+All rtPCR reactions utilized premade gelled format 96-well plates (Biotools, B &M Labs, S.A., Madrid, Spain), with the exception of blaTEM and mecA genes which used the Sybr GreenTM and TaqManTM probe, respectively.
+Our thermal cycle was the same for all the rtPCR reactions [6′ 95 °C, 40× (10″ 95 °C, 30″ 60 °C)], with alignment and extension in the same step, at constant temperature of 60 °C.
+A melting curve step was performed at the end of the qPCR reaction to validate the authenticity of the positive (Nieto-Claudin et al., 2019).
+We quantified the relative burden of each gene for each sample via the cycle threshold (Ct) for the 16S rRNA and the Ct value of the individual ARG using a previously published formula (Esperón et al., 2020).
 
 ## Data import and cleaning
-_Write code that reads in the file and cleans it so it's ready for analysis.
+
+*Write code that reads in the file and cleans it so it's ready for analysis.
 Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts.
@@
 as supplement in a paper).
-All materials, including files that contain code, should be commented well so everyone can follow along._
+All materials, including files that contain code, should be commented well so everyone can follow along.*
+
 ```{r}
@@
 
-raw <- read_csv(here('data/raw_data/StorkAMRraw_2020_2021.csv'))
-
-tibble(raw)
 ```
 
-## Statistical analysis
-_Basic summary statistics will be performed.
-Models will be built and fitted to data using the statistical package lme4 in Program R.
-We will evaluated model fit for all models using Akaike Information Criterion scores adjusted for small sample sizes.
-The best fit model was selected by having a ΔAICc < 2._
+### Statistical analysis
+
+Presence absence ARG results obtained from the fecal samples between 2020 to 2021 were used for simple summary statistics.
+Samples were classified as "multiresistant" if they were resistant to three or more of the 8 antibiotic classes that we evaluated for in this study (Blanco-Peña et al., 2017).
+In addition, we applied the following formula to estimate the percentage of bacteria harboring ARGs: x = 10\[2+0.33(ct16S-ctARG)\], where x individual percent gene burden in the sample (i.e., the estimated number of copies of the gene present per reaction).
+Results were expressed in log10, ranging from −8 ( zero to a negligible amount of the bacteria in the sample carried an ARG) to 2 (all or 100% of the bacteria in the sample carried an ARG).
+The inverse Log10 was then applied to results so they could be totalled and used to evalute total gene burdens accross sampling periods.
+
+Several linear mixed models (LMM) were constructed to evaluate multi-resistance and ARG burden as response variables with nest as a random factor.
+Covariates considered with each response variable included adult age, adult mean land fill use index (LUI), sample period (as described above), and nest success.
+Landfill use was quantified by physically observing a banded stork at Colmenar Viejo during weekly visits from March to June in 2021.
+The LUI was calculated as the number of observations of one particular bird within the total number of visits to landfill per year \<CITE Alejandro 2022\>.
+If a banded adult was not seen at the landfill during the breeding season, they were assigned a LUI of 0, suggestive of no resource provisioning at the landfill.
+All covariates were evaluated for correlation, no covariates were correlated with all the Spearman's correlation coefficients (rho) \< 0.5 and the p \> 0.05.
+All continuous variables (LUI, age, and nest success) were then standardized prior to analysis.
+
+All models were constructed with only 2021 data, as the COVID-19 pandemic prohibited the collection of LUI data in 2020.
+Models were built and fitted to data using the statistical package tidymodels in Program R (R version 4.2.1, www.r-project.org).
+
+#### (Note: still trying to figure out how to do this with the performance package) We evaluated model fit for all models using Akaike Information Criterion scores adjusted for small sample sizes (AICc: Burnham and Anderson, 2002).
+The best fit model was selected by having a ΔAICc \< 2.\_
 
@@
 
-_Use a combination of text/tables/figures to explore and describe your data.
+*Use a combination of text/tables/figures to explore and describe your data.
 Show the most important descriptive results here.
 Additional ones should go in the supplement.
-Even more can be in the R and Quarto files that are part of your project._
+Even more can be in the R and Quarto files that are part of your project.*
 
@@
 
-_To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
+*To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
 simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._
+Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*
 
@@
 #| echo: FALSE
-knitr::include_graphics(here("results","height_weight_stratified.png"))
+knitr::include_graphics(here("3 Analysis Scripts","LUIvsBurden_LME.tiff"))
+knitr::include_graphics(here("3 Analysis Scripts","LUIvsMDR_LME.tiff"))
 ```
@@
 
-_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
+*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
 This might again be code that is best placed in one or several separate R scripts that need to be well documented.
 You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here._
+Then you load them here.*
 
@@
 #| echo: FALSE
-resulttable2 = readRDS(here("results","resulttable2.rds"))
-knitr::kable(resulttable2)
+ModelAIC1 = readRDS(here("4 Results", "Tables", "ModelAIC1.rds"))
+knitr::kable(ModelAIC1)
+
+ModelAIC2 = readRDS(here("4 Results", "Tables", "ModelAIC2.rds"))
+knitr::kable(ModelAIC2)
 ```
@@
 ## Summary and Interpretation
-_Summarize what you did, what you found and what it means._
+
+-	As multi-drug resistance and class diversity increase (similar things I know) throughout the breeding season (compounding effect likely), mean LUI decreases and nest success decreases (makes sense because in your prior papers you have found increase nest success with increased LUI).
+
+-	Resistance gene burden appears to increase as mean LUI use and age increase. 
+o	Most of the burden is due to blaTEM, a common resistance gene associated with anthropogenic impact.
+Sampling period does not appear to explain burden, but the top blaTEM model did show a trend in burden increasing from the 2nd sampling period to the 3rd sampling period and age.
 
 ## Strengths and Limitations
-_Discuss what you perceive as strengths and limitations of your analysis._
+
+-	Not much statistically significant data, thus we may have to argue biological significance.
 
 ## Conclusions
-_What are the main take-home messages?_
 
-_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_
+LUI appears to be correlated with higher levels of AMR gene burden in storks.
+As LUI increases thorough the breeding season (Bialas et al 2021) resistance gene burden also increases with beta lactam resistance contributing to the majority of the burden.
+However, multidrug resistance appears to decrease as LUI increases, thus it is likely that storks are being exposed to antimicrobial resistance genes at other foraging areas (urban centers, agricultural pastures, etc.).
+Our results suggest that landfills may not be contributing significantly to the emergence and maintenance of AMR in this system.
+Little literature exists on the relationship between stork habitat selection and health outside of landfill use in Spain.
+Future efforts in stork AMR research should focus on exploring the relationship between agricultural land use and health.
+
+*What are the main take-home messages?*
+
+*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*
 

### 2023-03-20 5 Manuscript/manuscript/Manuscript.qmd
RF and SL troubleshooting session
Questions for Office Hours

@@
 
-*Study Area* — This study took place in Prado Herrero, a private cattle ranch located northwest of metropolitan Madrid and is surrounded by agriculture (e.g.
-beef cattle, cereal grains, legumes, and forage plants (“Renta Agraria,” 2019)).
+*Study Area* --- This study took place in Prado Herrero, a private cattle ranch located northwest of metropolitan Madrid and is surrounded by agriculture (e.g.
+beef cattle, cereal grains, legumes, and forage plants ("Renta Agraria," 2019)).
 Prado Herrero is located within a nationally protected area (Cuenca Alta del Manzanares Regional Park) and is just north of Santillana Reservoir.
 This reservoir that was declared an Important Bird Area by Regional Catalogue of Reservoirs and Wetlands of the Community of Madrid due to the numerous resident and migratory species that utilize this water source <CITE>.
-This cattle ranch has supported a productive white stork rookery where storks have been banded and monitored by biologists at UCM for over 20 years <CITE: Aguirre and Atienza>.
+This cattle ranch has supported a productive white stork rookery where storks have been banded and monitored by biologists at UCM for over 20 years \<CITE: Aguirre and Atienza\>.
 During the 2020 to 2021 breeding season, between the months of March to June, stork nests were identified, marked, and monitored for productivity.
@@
 Storks that breed within the Prado Herrero rookery are known to utilize Colmenar Viejo Landfill which is located approximately 12km southeast.
-Colmenar Viejo is an open-air landfill and it is second largest of it’s kind in the Madrid region <CITE: Alejandro 2022>.
+Colmenar Viejo is an open-air landfill and it is second largest of it's kind in the Madrid region \<CITE: Alejandro 2022\>.
 
-*Sample Collection* — Between March to May 2020 and 2021, we collected feces from marked nests with banded adults in known breeding pairs at 3 points of the breeding season; (1) an adult sample during incubation, (2) an early juvenile sample during the first two weeks of the chicks life when adults are believed to forage on natural sources  <CITE: >, and (3) a late juvenile sample after chicks were past two weeks of age when adults forage on anthropogenic resources.
+*Sample Collection* --- Between March to May 2020 and 2021, we collected feces from marked nests with banded adults in known breeding pairs at 3 points of the breeding season; (1) an adult sample during incubation, (2) an early juvenile sample during the first two weeks of the chicks life when adults are believed to forage on natural sources \<CITE: \>, and (3) a late juvenile sample after chicks were past two weeks of age when adults forage on anthropogenic resources.
 Nests were visited in the late mornings and approximately one gram of fresh feces was collected from the perimeter of the nest structure into a sterile Eppendorf tube ().
@@
 ### Molecular analysis of ARGs
-We performed total DNA extraction directly from fecal samples, by using a pressure filtration technique (QuickGene DNA Tissue Kit S, Fujifilm, Japan) following the manufacturer’s instructions.
+
+We performed total DNA extraction directly from fecal samples, by using a pressure filtration technique (QuickGene DNA Tissue Kit S, Fujifilm, Japan) following the manufacturer's instructions.
 The 16S rRNA gene was amplified in each DNA sample by real time PCR (rtPCR) in 10-fold dilutions of extracted samples, according to Jiang et al.
@@
 All rtPCR reactions utilized premade gelled format 96-well plates (Biotools, B &M Labs, S.A., Madrid, Spain), with the exception of blaTEM and mecA genes which used the Sybr GreenTM and TaqManTM probe, respectively.
-Our thermal cycle was the same for all the rtPCR reactions [6′ 95 °C, 40× (10″ 95 °C, 30″ 60 °C)], with alignment and extension in the same step, at constant temperature of 60 °C.
+Our thermal cycle was the same for all the rtPCR reactions \[6′ 95 °C, 40× (10″ 95 °C, 30″ 60 °C)\], with alignment and extension in the same step, at constant temperature of 60 °C.
 A melting curve step was performed at the end of the qPCR reaction to validate the authenticity of the positive (Nieto-Claudin et al., 2019).
@@
 
-## Data import and cleaning
+## Data import and cleaning (Reference file and give details HERE. Reference a ReadME/doc here)
 
@@
 
-## Exploratory/Descriptive analysis
+## Exploratory/Descriptive analysis (SUMMARY STATS IN EXPLORATION FILE TBD -\> Burden/MDR associated with each class. summary table? with actual summary stats)
 
@@
 #| echo: FALSE
-resulttable=readRDS("../../results/summarytable.rds")
-knitr::kable(resulttable)
+#resulttable=readRDS("../../results/summarytable.rds")
+#knitr::kable(resulttable)
 ```
@@
 ModelAIC1 = readRDS(here("4 Results", "Tables", "ModelAIC1.rds"))
+ModelAIC1
 knitr::kable(ModelAIC1)

### 2023-03-20 3 Analysis Scripts/StatiscalAnalysis_ModelFitting.qmd
RF and SL troubleshooting session
Questions for Office Hours
RF and SL AICc Troubleshooting

@@
 ```{r}
-AIC_table1 <- compare_performance(lmer_fit_null,lmer_fit_global, lmer_fit_1, lmer_fit_2, lmer_fit_3, lmer_fit_4, lmer_fit_5, lmer_fit_6, lmer_fit_7, lmer_fit_8, lmer_fit_9, lmer_fit_10, lmer_fit_11, lmer_fit_12)
+AIC_table1 <- compare_performance(lmer_fit_null,lmer_fit_global, lmer_fit_1, lmer_fit_2, lmer_fit_3, lmer_fit_4, lmer_fit_5, lmer_fit_6, lmer_fit_7, lmer_fit_8, lmer_fit_9, lmer_fit_10, lmer_fit_11, lmer_fit_12, rank = TRUE, metrics="AICc")
 
@@
 
-plot(compare_performance(lmer_fit_null,lmer_fit_global, lmer_fit_1, lmer_fit_2, lmer_fit_3, lmer_fit_4, lmer_fit_5, lmer_fit_6, lmer_fit_7, lmer_fit_8, lmer_fit_9, lmer_fit_10, lmer_fit_11, lmer_fit_12, rank = TRUE))
+plot(compare_performance(lmer_fit_null,lmer_fit_global, lmer_fit_1, lmer_fit_2, lmer_fit_3, lmer_fit_4, lmer_fit_5, lmer_fit_6, lmer_fit_7, lmer_fit_8, lmer_fit_9, lmer_fit_10, lmer_fit_11, lmer_fit_12))
 ```
@@
 ```{r}
-AIC_table2 <- compare_performance(lmer_fit2_null,lmer_fit2_global, lmer_fit2_1, lmer_fit2_2, lmer_fit2_3, lmer_fit2_4, lmer_fit2_5, lmer_fit2_6, lmer_fit2_7, lmer_fit2_8, lmer_fit2_9, lmer_fit2_10, lmer_fit2_11, lmer_fit2_12)
+AIC_table2 <- compare_performance(lmer_fit2_null, lmer_fit2_1, lmer_fit2_3, lmer_fit2_4, lmer_fit2_5, lmer_fit2_6, lmer_fit2_7, rank = TRUE, metrics="AICc")
 
@@
 
-plot(compare_performance(lmer_fit2_null,lmer_fit2_global, lmer_fit2_1, lmer_fit2_2, lmer_fit2_3, lmer_fit2_4, lmer_fit2_5, lmer_fit2_6, lmer_fit2_7, lmer_fit2_8, lmer_fit2_9, lmer_fit2_10, lmer_fit2_11, lmer_fit2_12, rank = TRUE))
+plot(compare_performance(lmer_fit2_null,lmer_fit2_5, lmer_fit2_1, lmer_fit2_3, lmer_fit2_4, lmer_fit2_6, lmer_fit2_7))
 ```
@@
 
-Questions: 
-Should over fit models be taken out of performance analysis?
-What is the best tool/package to control for small data sets?
+Questions: Should over fit models be taken out of performance analysis?
+
+\*\*What is the best tool/package to control for small data sets? If using weighted AIC is it appropriate to also use other methods of performance analysis (R2, RSME, etc).
+
+How to get Manuscript to popular rds and tiff files. GLMER vs LMER

### 2023-04-07 3 Analysis Scripts/StatiscalAnalysis_ModelFitting.qmd
Edits to full statistical analysis

--- 2023-03-20 3 Analysis Scripts/StatiscalAnalysis_ModelFitting.qmd
@@
 
-```{r}
+```{r, warning=FALSE, message=FALSE}
 #load needed packages. make sure they are installed.
@@
 library(performance) #evaluate model fit and performance
+library(see)
+library(cAIC4)
 
@@
 glmer_spec <- 
-  linear_reg() %>% 
+  logistic_reg() %>% 
   set_engine("glmer")
@@
 ```{r}
-lmer_fit_null <- 
-  lmer_spec %>% 
-  fit(MDR ~ (1|nes), data = data)
+glmer_fit_null <- 
+  glmer_spec %>% 
+  fit(as.factor(MDR) ~ (1|nes), data = data)
 
-lmer_fit_global <- 
-  lmer_spec %>% 
-  fit(MDR ~ s.lui + s.age + samp + s.nsuccess +(1|nes), data = data)
+glmer_fit_global <- 
+  glmer_spec %>% 
+  fit(as.factor(MDR) ~ s.lui + s.age + samp + s.nsuccess +(1|nes), data = data)
 
 #Univariate models
-lmer_fit_1 <- 
-  lmer_spec %>% 
-  fit(MDR ~ s.lui +(1|nes), data = data)
+glmer_fit_1 <- 
+  glmer_spec %>% 
+  fit(as.factor(MDR) ~ s.lui +(1|nes), data = data)
 
-lmer_fit_2 <- 
-  lmer_spec %>% 
-  fit(MDR ~ s.age  +(1|nes), data = data)
+glmer_fit_2 <- 
+  glmer_spec %>% 
+  fit(as.factor(MDR) ~ s.age  +(1|nes), data = data)
 
-lmer_fit_3 <- 
-  lmer_spec %>% 
-  fit(MDR ~ samp +(1|nes), data = data)
+glmer_fit_3 <- 
+  glmer_spec %>% 
+  fit(as.factor(MDR) ~ samp +(1|nes), data = data)
 
-lmer_fit_4 <- 
-  lmer_spec %>% 
-  fit(MDR ~ s.nsuccess +(1|nes), data = data)
+glmer_fit_4 <- 
+  glmer_spec %>% 
+  fit(as.factor(MDR) ~ s.nsuccess +(1|nes), data = data)
 
-lmer_fit_5 <- 
-  lmer_spec %>% 
-  fit(MDR ~ s.lui + samp +(1|nes), data = data)
+#Multivariate models
 
-lmer_fit_6 <- 
-  lmer_spec %>% 
-  fit(MDR ~ s.lui + s.nsuccess +(1|nes), data = data)
+glmer_fit_5 <- 
+  glmer_spec %>% 
+  fit(as.factor(MDR) ~ s.lui + samp +(1|nes), data = data)
 
-lmer_fit_7 <- 
-  lmer_spec %>% 
-  fit(MDR ~ samp + s.nsuccess +(1|nes), data = data)
+glmer_fit_6 <- 
+  glmer_spec %>% 
+  fit(as.factor(MDR) ~ s.lui + s.nsuccess +(1|nes), data = data)
 
-lmer_fit_8 <- 
-  lmer_spec %>% 
-  fit(MDR ~ s.lui*s.age + samp + s.nsuccess +(1|nes), data = data)
+glmer_fit_7 <- 
+  glmer_spec %>% 
+  fit(as.factor(MDR) ~ samp + s.nsuccess +(1|nes), data = data)
 
-lmer_fit_9 <- 
-  lmer_spec %>% 
-  fit(MDR ~ s.lui + s.age*samp + s.nsuccess +(1|nes), data = data)
+glmer_fit_8 <- 
+  glmer_spec %>% 
+  fit(as.factor(MDR) ~ s.lui*s.age + samp + s.nsuccess +(1|nes), data = data)
 
-lmer_fit_10 <- 
-  lmer_spec %>% 
-  fit(MDR ~ s.lui + s.age + samp*s.nsuccess +(1|nes), data = data)
+glmer_fit_9 <- 
+  glmer_spec %>% 
+  fit(as.factor(MDR) ~ s.lui + s.age*samp + s.nsuccess +(1|nes), data = data)
 
-lmer_fit_11 <- 
-  lmer_spec %>% 
-  fit(MDR ~ s.lui + s.age +(1|nes), data = data)
+glmer_fit_10 <- 
+  glmer_spec %>% 
+  fit(as.factor(MDR) ~ s.lui + s.age + samp*s.nsuccess +(1|nes), data = data)
 
-lmer_fit_12 <- 
-  lmer_spec %>% 
-  fit(MDR ~ s.lui + s.age +  s.nsuccess +(1|nes), data = data)
+glmer_fit_11 <- 
+  glmer_spec %>% 
+  fit(as.factor(MDR) ~ s.lui + s.age +(1|nes), data = data)
+
+glmer_fit_12 <- 
+  glmer_spec %>% 
+  fit(as.factor(MDR) ~ s.lui + s.age +  s.nsuccess +(1|nes), data = data)
 ```
@@
 ```{r}
-AIC_table1 <- compare_performance(lmer_fit_null,lmer_fit_global, lmer_fit_1, lmer_fit_2, lmer_fit_3, lmer_fit_4, lmer_fit_5, lmer_fit_6, lmer_fit_7, lmer_fit_8, lmer_fit_9, lmer_fit_10, lmer_fit_11, lmer_fit_12)
+AIC_table1 <- compare_performance(glmer_fit_null,glmer_fit_global, glmer_fit_1, glmer_fit_2, glmer_fit_3, glmer_fit_4, glmer_fit_5, glmer_fit_6, glmer_fit_7, glmer_fit_8, glmer_fit_9, glmer_fit_10, glmer_fit_11, glmer_fit_12, rank = TRUE, metrics="AICc")
 
 print(AIC_table1)
-AIC_table1 = here("4 Results","Tables", "ModelAIC1.rds")
-saveRDS(AIC_table1, file = AIC_table1)
 
-plot(compare_performance(lmer_fit_null,lmer_fit_global, lmer_fit_1, lmer_fit_2, lmer_fit_3, lmer_fit_4, lmer_fit_5, lmer_fit_6, lmer_fit_7, lmer_fit_8, lmer_fit_9, lmer_fit_10, lmer_fit_11, lmer_fit_12, rank = TRUE))
+AIC_table1 %>%
+  saveRDS(here("4 Results","Tables", "ModelAIC1.rds"))
+
+plot(compare_performance(glmer_fit_null,glmer_fit_global, glmer_fit_1, glmer_fit_2, glmer_fit_3, glmer_fit_4, glmer_fit_5, glmer_fit_6, glmer_fit_7, glmer_fit_8, glmer_fit_9, glmer_fit_10, glmer_fit_11, glmer_fit_12))
 ```
@@
 ```{r}
-check_model(lmer_fit_11)
+check_model(glmer_fit_11)
 
@@
 
-model1 <- broom::tidy(lmer_fit_11)
+model1 <- broom::tidy(glmer_fit_11)
 print(model1)
@@
 # save fit results table  
-model1 = here("4 Results", "Models", "LMERmodel1.rds")
-saveRDS(model1, file = model1)
+model1 %>%
+  saveRDS(here("4 Results", "Models", "LMERmodel1.rds"))
 
-MLp <- plot_model(lmer_fit_11,terms = 's.lui', type="pred", 
+MLp <- plot_model(glmer_fit_11,terms = 's.lui', type="pred", 
            show.data = TRUE, show.values = TRUE, show.p = TRUE,
@@
 ```{r}
-AIC_table2 <- compare_performance(lmer_fit2_null,lmer_fit2_global, lmer_fit2_1, lmer_fit2_2, lmer_fit2_3, lmer_fit2_4, lmer_fit2_5, lmer_fit2_6, lmer_fit2_7, lmer_fit2_8, lmer_fit2_9, lmer_fit2_10, lmer_fit2_11, lmer_fit2_12)
+AIC_table2 <- compare_performance(lmer_fit2_null, lmer_fit2_global, lmer_fit2_1, lmer_fit2_2, lmer_fit2_3, lmer_fit2_4, lmer_fit2_5, lmer_fit2_6, lmer_fit2_7, lmer_fit2_8,
+                                  lmer_fit2_9, lmer_fit2_10, lmer_fit2_11, lmer_fit2_12, rank = TRUE, metrics="AICc")
 
 print(AIC_table2)
-AIC_table2 = here("4 Results","Tables", "ModelAIC2.rds")
-saveRDS(AIC_table2, file = AIC_table2)
 
-plot(compare_performance(lmer_fit2_null,lmer_fit2_global, lmer_fit2_1, lmer_fit2_2, lmer_fit2_3, lmer_fit2_4, lmer_fit2_5, lmer_fit2_6, lmer_fit2_7, lmer_fit2_8, lmer_fit2_9, lmer_fit2_10, lmer_fit2_11, lmer_fit2_12, rank = TRUE))
+AIC_table2 %>%
+ saveRDS(here("4 Results","Tables", "ModelAIC2.rds"))
+
+plot(compare_performance(lmer_fit2_null, lmer_fit2_global, lmer_fit2_1, lmer_fit2_2, lmer_fit2_3, lmer_fit2_4, lmer_fit2_5, lmer_fit2_6, lmer_fit2_7, lmer_fit2_8,
+                                  lmer_fit2_9, lmer_fit2_10, lmer_fit2_11, lmer_fit2_12))
 ```
@@
 # save fit results table  
-model2 = here("4 Results","Models", "LMERmodel2.rds")
-saveRDS(model2, file = model2)
+model2 %>% 
+  saveRDS(here("4 Results","Models", "LMERmodel2.rds"))
 
@@
 ```
-
-Preliminary results
-
-Questions: 
-Should over fit models be taken out of performance analysis?
-What is the best tool/package to control for small data sets?

### 2023-04-07 3 Analysis Scripts/StatiscalAnalysis_ModelFitting.qmd
Updates to manuscript and statistical analysis

@@
 model1 %>%
-  saveRDS(here("4 Results", "Models", "LMERmodel1.rds"))
+  saveRDS(here("4 Results", "Models", "GLMERmodel1.rds"))
 
@@
 
-model2 <- broom::tidy(lmer_fit2_5)
+model2 <- broom::tidy(lmer_fit2_global)
 print(model2)
@@
 
-BLp <- plot_model(lmer_fit2_5,terms = 's.lui', type="pred", 
+BLp <- plot_model(lmer_fit2_global,terms = 's.lui', type="pred", 
            show.data = TRUE, show.values = TRUE, show.p = TRUE,

### 2023-04-07 5 Manuscript/manuscript/Manuscript.qmd
Edits to full statistical analysis
Updates to manuscript and statistical analysis

@@
 
-```{r, echo=FALSE, message=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 # load a few R packages
@@
 
-```{r}
+```{r, warning=FALSE}
 library(tidyverse)
@@
 Results were expressed in log10, ranging from −8 ( zero to a negligible amount of the bacteria in the sample carried an ARG) to 2 (all or 100% of the bacteria in the sample carried an ARG).
-The inverse Log10 was then applied to results so they could be totalled and used to evalute total gene burdens accross sampling periods.
+The inverse Log10 was then applied to results so they could be totaled and used to evaluate total gene burdens across sampling periods.
 
@@
 
-@tbl-summarytable shows a summary of the data.
-
-Note the loading of the data providing a **relative** path using the `../../` notation.
-(Two dots means a folder up).
-You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path.
-You can also use the `here` R package to create paths.
-See examples of that below.
-
-```{r}
-#| label: tbl-summarytable
-#| tbl-cap: "Data summary table."
-#| echo: FALSE
-#resulttable=readRDS("../../results/summarytable.rds")
-#knitr::kable(resulttable)
-```
-
 ## Basic statistical analysis
@@
 
-@fig-result shows a scatterplot figure produced by one of the R scripts.
-
 ```{r}
-#| label: fig-result
-#| fig-cap: "Height and weight stratified by gender."
+#| label: Figure1
+#| fig-cap: "Total antimicrobial gene burden as predicted by the variables landfill use index, age, sampling period, and nest success."
 #| echo: FALSE
 knitr::include_graphics(here("3 Analysis Scripts","LUIvsBurden_LME.tiff"))
+
+#| label: Figure2
+#| fig-cap: "Multi-drug resistance as predicted by the variables landfill use index and age."
+#| echo: FALSE
 knitr::include_graphics(here("3 Analysis Scripts","LUIvsMDR_LME.tiff"))
@@
 
-Example @tbl-resulttable2 shows a summary of a linear model fit.
+Binomial generalized linear mixed models (GLMMs) were used to predict multi-drug resistance (MDR) in White Storks, with nest identification classified as a random effects parameter.
+Of these, the highest performing model contained the predictors landfill use index (LUI) and age (@Table1).
+In this model, an increase in the age of the bird was associated with a higher likelihood of MDR being present (@Table2), while an increase in LUI was found to be associated with a decreased likelihood of MDR presence.
+A second competitive model, with a 97.20% performance score, contained only age as the predictor.
+
+Linear mixed models (LMMs) were used to predict total antimicrobial gene burden in White Storks, with nest identification again classified as a random effects parameter.
+The global model, containing the variables landfill use index, age, sample period, and nest success, was the highest performing model (@Table3).
+In this model, each variable was positively correlated with antimicrobial gene burden (@Table4).
+The next highest performing model contained age as a single predictor, with a model performance score of 86.39%.
 
 ```{r}
-#| label: tbl-resulttable2
-#| tbl-cap: "Linear model fit table."
+#| label: Table1
+#| tbl-cap: "Small sample adjusted Akaike's Information Criteria (AICc) model weights and performance scores for binomial generalized linear mixed models predicting multi-drug resistance in White Storks (Ciconia ciconia) in Madrid, Spain.
+"
 #| echo: FALSE
 ModelAIC1 = readRDS(here("4 Results", "Tables", "ModelAIC1.rds"))
-ModelAIC1
 knitr::kable(ModelAIC1)
 
+#| label: Table2
+#| tbl-cap: "Parameter estimates for fixed and random effects parameters and standard error of the highest performing model predicting multi-drug resistance in White Storks in Madrid, Spain."
+#| echo: FALSE
+GLMERmodel1 = readRDS(here("4 Results", "Tables", "GLMERmodel1.rds"))
+knitr::kable(GLMERmodel1)
+
+```
+
+```{r}
+#| label: Table3
+#| tbl-cap: "Small sample adjusted Akaike's Information Criteria (AICc) model weights and performance scores for linear mixed models predicting total antimicrobial gene burden in White Storks (Ciconia ciconia) in Madrid, Spain.
+"
+#| echo: FALSE
 ModelAIC2 = readRDS(here("4 Results", "Tables", "ModelAIC2.rds"))
 knitr::kable(ModelAIC2)
+
+#| label: Table4
+#| tbl-cap: "Parameter estimates for fixed and random effects parameters and standard error of the highest performing model predicting total antimicrobial gene burden in White Storks in Madrid, Spain."
+#| echo: FALSE
+GLMERmodel1 = readRDS(here("4 Results", "Tables", "LMERmodel2.rds"))
+knitr::kable(GLMERmodel1)
+
 ```

### 2023-04-07 5 Manuscript/manuscript/Manuscript.qmd
readme.md Files + Background
Missing bibTEX files (Currently troubleshooting)

--- 2023-03-20 5 Manuscript/manuscript/Manuscript.qmd
@@
 library(knitr)
+library(knitcitations)
+library(citr)
+library(devtools)
+cleanbib()
 ```
@@
 
-## General Background Information
+## General Background Information (WILL FIGHT WITH BIBTEX TOMORROW)
 
-*Provide enough background on your topic that others can understand the why and how of your analysis*
+Antimicrobial resistance (AMR) in wildlife was recently highlighted as being a critical research need (Dolejska & Literak, 2019).
+While AMR is accepted as a global human concern, it has only recently been utilized to evaluated wildlife health and anthropogenic impacts on environments.
+The majority of resistance research in wildlife has been targeted towards birds due to their interinsict life history traits such as long migrations and the ability to fly between environments <CITE>.
+However, the importance of birds in respect to the epidemiology of AMR remains poorly understood (Radhouani et al., 2014).
+A growing body of literature is trying to understand the role free-roaming birds play in the emergence, maintenance, and transmission of AMR upon the landscape.
+
+Birds have a long history of being used as indicators for environmental and human health.
+For example, wading bird surveillance is commonplace in south Florida to monitor heavy metal levels like mercury in the greater everglades ecosystem <CITE>.
+Birds are also often utilized in zoonotic pathogen surveillance (i.e., USDA sentinel chickens to monitor West Nile Virus) < >.
+In recent years, a similar approach has been utilized to study AMR prevalence/impacts on different landscapes <>.
+In 2019, the USGS revealed that more than half of GPS-tracked gulls (Larus ssp.) had acquired antimicrobial resistant E.
+coli from landfills and disseminated it long distances to pristine, unaltered habitats (Ahlstrom et al., 2019).
+This supports the belief that birds who utilize and move between human-dominated and natural landscapes may become both reservoirs and efficient transporters of resistance determinants to natural environments.For wildlife in general, resistance prevalence tends to increase with proximity to human populations (Niu et al., 2013).
+It is believed that animals that rely on anthropogenic resources (e.g., landfills) are often concentrated at high densities that promote the large-scale mixing of bacteria, which facilitate the rise of novel resistant strains (Wellington et al., 2013).
+Moreover, ecological factors such as migration and high densities during breeding season may increase acquisition rates and the dissemination of AMR <CITE>.
+
+Understanding the prevalence and movement of AMR is important to not only humans and environmental health, but also wildlife health.
+Currently, the carriage of AMR genes presents an unknown risk to the individual birds but has larger implications for free-living avian communities.
+Resistance genes are thought to alter the fitness of a pathogen (Friedman et al., 2016), and have been correlated with increased virulence (Escudeiro et al., 2019), in turn increasing morbidity and mortality in affected hosts.
+Finally, at the population management level, birds have also been held ‘responsible’ for the dissemination of enteric pathogens (e.g., E.
+coli, Salmonella spp.) that have caused outbreaks in produce, yet, they are only moving these bacteria from areas already contaminated by human activities (Hoelzer et al., 2011).
+Thus, understanding how avian behaviors that overlap with risky environmental factors, (e.g., foraging in agricultural areas), impact the carriage of resistance, is key because: 1) it allows us to understand the exact role birds play in AMR dissemination, and  2) it informs targeted population management and conservation strategies—such as discouraging birds from utilizing certain areas.
+
+The western white stork population provides the unique opportunity to explore how anthropogenic land use influences the carriage of antimicrobial resistance genes.
+White storks (Ciconia ciconia) were historically trans-Sahara migrants whose ecology has dramatically changed in response to anthropogenic pressure.
+Since the 1990s, Spanish white stork populations have steadily increased, due to their ability to exploit anthropogenic habitats, especially in areas with abundant resources especially urban open landfills (Cramp and (ed), 1985; Tortosa et al., 2002).
+Currently, colonies in Madrid, Spain with increased proximity to landfills have improved breeding outcomes and nestling quality, however eggs and nestlings also have increased pollutant loads (Vergara et al., 2006) (Jiang et al., 2013),(Blázquez et al., 2006; Tortosa et al., 2002).
+This has in turn affected their natural history as subsets of the population begun to shorten or abandon their winter migration (Gordo et al., 2007; Marchamalo et al., 1997; Reed et al., 2003; Tortosa et al., 1995; Vergara et al., 2004).
+Although the conservation status of the white stork is of Least Concern in Spain, it has disappeared from large areas of its historical range in Europe and thus, the populations in Portugal and Spain are of greater significance.
+
+Most studies on AMR surveillance in wildlife are performed by culture‐dependent methods, (e.g., the isolation of a specific pathogen such as Escherichia coli) as indicators of AMR.
+However, since most bacteria are not culturable, the detection of AMR using traditional methods might not be representative of the whole resistant microbiota, sometimes referred to as the resistome .
+Real time PCR (rtPCR), a technique not dependent on culture, offers that ability to quantify the presence and abundance of antimicrobial resistance genes (ARGs),  in white storks.
+This technique has recently been used in three studies to investigate the presence and amount of ARGs in the environment (e.g., soil and manure) and wildlife (e.g., Galapagos tortoise, gulls, guia) (Esperón et al., 2018; Nieto-Claudin et al., 2019; Sacristán et al., 2020).
+Add blurb about HGT… For instance, birds that encounter human waste (e.g., by drinking contaminated water) can acquire resistance genes through the exchange of naturally present genes in the bacteria they carry, with the genes in the waste—this can co-select for mobile genetic elements carrying multiple resistant genes (Wellington et al., 2013).
+
+<Insert Paragraph with objectives>
 

### 2023-04-08 5 Manuscript/manuscript/Manuscript.qmd
BibTEX added

--- 2023-03-20 5 Manuscript/manuscript/Manuscript.qmd
@@
     highlight-style: github
-bibliography: ../dataanalysis_template_references.bib
-csl: ../apa.csl
+bibliography: EW_AMR.bib
 ---
 
-```{r, echo=FALSE, message=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 # load a few R packages
@@
 library(knitr)
+library(knitcitations)
 ```
@@
 
-## General Background Information
+## General Background Information (WILL FIGHT WITH BIBTEX TOMORROW)
 
-*Provide enough background on your topic that others can understand the why and how of your analysis*
+Antimicrobial resistance (AMR) in wildlife was recently highlighted as being a critical research need [@Dolejska2019].
+While AMR is accepted as a global human health concern, it has only recently been utilized to evaluated wildlife health and anthropogenic impacts on environments (i.e., environmental health).
+The majority of resistance research in wildlife has been targeted towards birds, however, the importance of birds in respect to the epidemiology of AMR remains poorly understood [@Radhouani2014].
+Now, a growing body of literature is trying to understand the role free-roaming birds play in the emergence, maintenance, and transmission of AMR upon the landscape.
 
-## Description of data and data source
+Birds have a long history of being used as indicators for environmental and human health.
+For example, wading bird surveillance is commonplace in south Florida to monitor heavy metal levels in the greater everglades ecosystem.
+Birds are also often utilized in zoonotic pathogen surveillance (i.e., USDA sentinel chickens to monitor West Nile Virus).
+In recent years, a similar approach has been utilized to study AMR prevalence/impacts on different landscapes.
+Currently, the carriage of AMR genes presents an unknown risk to the individual birds and may have greater implications for free-living avian communities and conservation.
+Resistance genes are thought to alter the fitness of a pathogen [@Friedman2016], and have been correlated with increased virulence [@Escudeiro2019], in turn increasing morbidity and mortality in affected hosts.
+Finally, at the population management level, birds have also been held ‘responsible’ for the dissemination of enteric pathogens (e.g., *Escherichia coli*, *Salmonella* spp.) that have caused outbreaks in produce, yet, they are only moving these bacteria from areas already contaminated by human activities [@Hoelzer2011].
+Thus, understanding how avian behaviors that overlap with risky environmental factors (e.g., foraging in agricultural areas or landfills) impact the carriage of resistance, is key because: 1) it allows us to better understand the role birds play in AMR dissemination, and  2) it informs targeted population management and conservation strategies—such as discouraging birds from utilizing certain areas.
 
-*We have 126 observations taken over a period of 2 years.
-We are evaluating for the presence and burden of antimicrobial resistance genes in white stork feces.
-Codebook is a WIP located in raw data folder.*
+In 2019, the USGS revealed that more than half of GPS-tracked gulls (*Larus* ssp.) had acquired antimicrobial resistant *E.
+coli* from landfills and disseminated it long distances to pristine, unaltered habitats [@Ahlstrom2019].
+This supports the belief that birds who utilize and move between human-dominated and natural landscapes may become both reservoirs and efficient transporters of resistance determinants to natural environments.For wildlife in general, resistance prevalence tends to increase with proximity to human populations [@Niu2013].
+It is believed that animals that rely on anthropogenic resources are often concentrated at high densities that promote the large-scale mixing of bacteria, encouranging horizontal gene transfer (HGT) facilitating the rise of novel resistant strains [@Wellington2013].
+For instance, birds that encounter human waste (e.g., by drinking contaminated water) can acquire resistance genes through the exchange of naturally present genes in the bacteria they carry, with the genes in the waste—this can co-select for mobile genetic elements carrying multiple resistant genes [@Wellington2013].
+Moreover, ecological factors such as migration and high densities during breeding season may increase acquisition rates and the dissemination of AMR .
+The western white stork are one such species that heavily utilizes agricultural and urbam areas, providing a unique opportunity to explore how anthropogenic land use influences the carriage of antimicrobial resistance genes.
+
+White storks (*Ciconia ciconia*) were historically trans-Sahara migrants whose ecology has dramatically changed in response to anthropogenic pressure.
+Since the 1990s, Spanish white stork populations have steadily increased, due to their ability to exploit anthropogenic habitats, especially in areas with abundant resources especially open landfills [@Cramp1985; @Tortosa2002].
+Currently, colonies in Madrid, Spain with increased proximity to landfills have improved breeding outcomes and nestling quality, however eggs and nestlings also have increased pollutant loads [@Vergara2006; @Jiang2013; @Saez2008; @Tortosa2002].
+This has in turn affected their natural history.
+Subsets of the population have begun to shorten or abandon their winter migration [@Gordo2007; @Tortosa1995; @Vergara2004].
+Although the conservation status of the white stork is of Least Concern in Spain, it has disappeared from large areas of its historical range in Europe and thus, the populations in Portugal and Spain are is integral to the species conservation.
+
+Most studies on AMR surveillance in wildlife are performed by culture‐dependent methods, (e.g., the isolation of a specific pathogen such as *E.
+coli*) as indicators of AMR.
+However, since most bacteria are not culturable, the detection of AMR using traditional methods might not be representative of the whole resistant microbiota, sometimes referred to as the resistome [@Surette2017].
+Real time PCR (rtPCR), a technique not dependent on culture, offers that ability to quantify the presence and abundance of antimicrobial resistance genes (ARGs), in white storks.
+This technique has recently been used in three studies to investigate the presence and amount of ARGs in the environment (e.g., soil and manure) and wildlife (e.g., Galapagos tortoise, gulls, guignas) [@Esperon2014; @NietoClaudin2019; @Sacristan2020].
+In this study we will explore the ARG load and diversity in a population of white stork in Madrid Spain during the 2020 and 2021.
 
@@
 
-*To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key.
-Then you can include like this:*
-
-Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a]
-
 {{< pagebreak >}}
@@
 
+### Description of data and data source
+
+We have 126 observations taken over a period of 2 years.
+We are evaluating for the presence and burden of antimicrobial resistance genes in white stork feces.
+Codebook is a WIP located in raw data folder.
+
 ### Experimental Design
@@
 *Study Area* --- This study took place in Prado Herrero, a private cattle ranch located northwest of metropolitan Madrid and is surrounded by agriculture (e.g.
-beef cattle, cereal grains, legumes, and forage plants ("Renta Agraria," 2019)).
+beef cattle, cereal grains, legumes, and forage plants.
 Prado Herrero is located within a nationally protected area (Cuenca Alta del Manzanares Regional Park) and is just north of Santillana Reservoir.
-This reservoir that was declared an Important Bird Area by Regional Catalogue of Reservoirs and Wetlands of the Community of Madrid due to the numerous resident and migratory species that utilize this water source <CITE>.
-This cattle ranch has supported a productive white stork rookery where storks have been banded and monitored by biologists at UCM for over 20 years \<CITE: Aguirre and Atienza\>.
+This reservoir that was declared an Important Bird Area by Regional Catalogue of Reservoirs and Wetlands of the Community of Madrid due to the numerous resident and migratory species that utilize this water source.
+This cattle ranch has supported a productive white stork rookery where storks have been banded and monitored by biologists at UCM for over 20 years [@Aguirre2009].
 During the 2020 to 2021 breeding season, between the months of March to June, stork nests were identified, marked, and monitored for productivity.
 Of marked nests between the 2020 and 2021 breeding seasons, 31 with banded adults were used both years to lay eggs successfully.
-All 31 nests were located in ash trees (Fraxinus angustifolius ) found within the cattle pasture.
+All 31 nests were located in ash trees (*Fraxinus angustifolius*) found within the cattle pasture.
 Storks that breed within the Prado Herrero rookery are known to utilize Colmenar Viejo Landfill which is located approximately 12km southeast.
-Colmenar Viejo is an open-air landfill and it is second largest of it's kind in the Madrid region \<CITE: Alejandro 2022\>.
+Colmenar Viejo is an open-air landfill and it is second largest of it's kind in the Madrid region [@Lopez-Garcia2021].
 
-*Sample Collection* --- Between March to May 2020 and 2021, we collected feces from marked nests with banded adults in known breeding pairs at 3 points of the breeding season; (1) an adult sample during incubation, (2) an early juvenile sample during the first two weeks of the chicks life when adults are believed to forage on natural sources \<CITE: \>, and (3) a late juvenile sample after chicks were past two weeks of age when adults forage on anthropogenic resources.
-Nests were visited in the late mornings and approximately one gram of fresh feces was collected from the perimeter of the nest structure into a sterile Eppendorf tube ().
+*Sample Collection* --- Between March to May 2020 and 2021, we collected feces from marked nests with banded adults in known breeding pairs at 3 points of the breeding season; (1) an adult sample during incubation, (2) an early juvenile sample during the first two weeks of the chicks life when adults are believed to forage on natural sources, and (3) a late juvenile sample after chicks were past two weeks of age when adults forage on anthropogenic resources.
+Nests were visited in the late mornings and approximately one gram of fresh feces was collected from the perimeter of the nest structure into a sterile Eppendorf tube.
 Samples were maintained cold in a portable cooler with frozen gel packs and frozen in a −20°C freezer within 4 hours of collection and processed at a later date.
@@
 We performed total DNA extraction directly from fecal samples, by using a pressure filtration technique (QuickGene DNA Tissue Kit S, Fujifilm, Japan) following the manufacturer's instructions.
-The 16S rRNA gene was amplified in each DNA sample by real time PCR (rtPCR) in 10-fold dilutions of extracted samples, according to Jiang et al.
-(2013)<CITE:Fix>.
-A DNA sample was considered validated when a ten-fold dilution showed a cycle threshold (Ct) less than 25 (Esperón et al., 2018).
+The 16S rRNA gene was amplified in each DNA sample by real time PCR (rtPCR) in 10-fold dilutions of extracted samples, according to @Jiang2013.
+A DNA sample was considered validated when a ten-fold dilution showed a cycle threshold (Ct) less than 25 [@Esperon2020].
 Once validated, we analyzed DNA samples by with a panel of 21 different ARGs encoding resistance to eight different antimicrobial classes: tetracyclines (tet(A), tet(B), tet(Y), tet(K), tet(M), tet(Q), tet(S), and tet(W)), sulfonamides (sulI and sulII), aminoglycosides (str and aadA), phenicols (catI and catII), macrolides (ermB and ermF), quinolones (qnrS and qnrB), betalactams (blaTEM and mecA), and polymyxins (mcr-1).
@@
 Our thermal cycle was the same for all the rtPCR reactions \[6′ 95 °C, 40× (10″ 95 °C, 30″ 60 °C)\], with alignment and extension in the same step, at constant temperature of 60 °C.
-A melting curve step was performed at the end of the qPCR reaction to validate the authenticity of the positive (Nieto-Claudin et al., 2019).
-We quantified the relative burden of each gene for each sample via the cycle threshold (Ct) for the 16S rRNA and the Ct value of the individual ARG using a previously published formula (Esperón et al., 2020).
+A melting curve step was performed at the end of the qPCR reaction to validate the authenticity of the positive [@NietoClaudin2019].
+We quantified the relative burden of each gene for each sample via the cycle threshold (Ct) for the 16S rRNA and the Ct value of the individual ARG using a previously published formula in @Esperon2020.
 
@@
 Presence absence ARG results obtained from the fecal samples between 2020 to 2021 were used for simple summary statistics.
-Samples were classified as "multiresistant" if they were resistant to three or more of the 8 antibiotic classes that we evaluated for in this study (Blanco-Peña et al., 2017).
+Samples were classified as "multiresistant" if they were resistant to three or more of the 8 antibiotic classes that we evaluated for in this study [@Blanco-Pena2017].
 In addition, we applied the following formula to estimate the percentage of bacteria harboring ARGs: x = 10\[2+0.33(ct16S-ctARG)\], where x individual percent gene burden in the sample (i.e., the estimated number of copies of the gene present per reaction).
 Results were expressed in log10, ranging from −8 ( zero to a negligible amount of the bacteria in the sample carried an ARG) to 2 (all or 100% of the bacteria in the sample carried an ARG).
-The inverse Log10 was then applied to results so they could be totalled and used to evalute total gene burdens accross sampling periods.
+The inverse Log10 was then applied to results so they could be totaled and used to evaluate total gene burdens across sampling periods.
 
@@
 Landfill use was quantified by physically observing a banded stork at Colmenar Viejo during weekly visits from March to June in 2021.
-The LUI was calculated as the number of observations of one particular bird within the total number of visits to landfill per year \<CITE Alejandro 2022\>.
+The LUI was calculated as the number of observations of one particular bird within the total number of visits to landfill per year [@Lopez-Garcia2021].
 If a banded adult was not seen at the landfill during the breeding season, they were assigned a LUI of 0, suggestive of no resource provisioning at the landfill.
@@
 #| label: fig-result
-#| fig-cap: "Height and weight stratified by gender."
+#| fig-cap: "Linear Mixed Model Output with 95% CI."
 #| echo: FALSE
@@
 #| label: tbl-resulttable2
-#| tbl-cap: "Linear model fit table."
+#| tbl-cap: "Model AICc Tables."
 #| echo: FALSE
@@
 LUI appears to be correlated with higher levels of AMR gene burden in storks.
-As LUI increases thorough the breeding season (Bialas et al 2021) resistance gene burden also increases with beta lactam resistance contributing to the majority of the burden.
+As LUI increases thorough the breeding season [@Bialas2020] resistance gene burden also increases with beta lactam resistance contributing to the majority of the burden.
 However, multidrug resistance appears to decrease as LUI increases, thus it is likely that storks are being exposed to antimicrobial resistance genes at other foraging areas (urban centers, agricultural pastures, etc.).
@@
 
-*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*
-
-This paper [@leek2015] discusses types of analyses.
-
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
-
-Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header.
-Many more style files for almost any journal [are available](https://www.zotero.org/styles).
-You also specify the location of your bibtex reference file in the YAML.
-You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.
-
 {{< pagebreak >}}

### 2023-04-10 5 Manuscript/manuscript/Manuscript.qmd

@@
 
-@fig-result shows a scatterplot figure produced by one of the R scripts.
-
 ```{r}
@@
 knitr::include_graphics(here("3 Analysis Scripts","LUIvsBurden_LME.tiff"))
+
+#| label: Figure2
+#| fig-cap: "Multi-drug resistance as predicted by the variables landfill use index and age."
+#| echo: FALSE
 knitr::include_graphics(here("3 Analysis Scripts","LUIvsMDR_LME.tiff"))
@@
 
-Example @tbl-resulttable2 shows a summary of a linear model fit.
+Binomial generalized linear mixed models (GLMMs) were used to predict multi-drug resistance (MDR) in White Storks, with nest identification classified as a random effects parameter.
+Of these, the highest performing model contained the predictors landfill use index (LUI) and age (@Table1).
+In this model, an increase in the age of the bird was associated with a higher likelihood of MDR being present (@Table2), while an increase in LUI was found to be associated with a decreased likelihood of MDR presence.
+A second competitive model, with a 97.20% performance score, contained only age as the predictor.
+
+Linear mixed models (LMMs) were used to predict total antimicrobial gene burden in White Storks, with nest identification again classified as a random effects parameter.
+The global model, containing the variables landfill use index, age, sample period, and nest success, was the highest performing model (@Table3).
+In this model, each variable was positively correlated with antimicrobial gene burden (@Table4).
+The next highest performing model contained age as a single predictor, with a model performance score of 86.39%.
 
@@
 ModelAIC1 = readRDS(here("4 Results", "Tables", "ModelAIC1.rds"))
-ModelAIC1
 knitr::kable(ModelAIC1)
 
+#| label: Table2
+#| tbl-cap: "Parameter estimates for fixed and random effects parameters and standard error of the highest performing model predicting multi-drug resistance in White Storks in Madrid, Spain."
+#| echo: FALSE
+GLMERmodel1 = readRDS(here("4 Results", "Tables", "GLMERmodel1.rds"))
+knitr::kable(GLMERmodel1)
+
+```
+
+```{r}
+#| label: Table3
+#| tbl-cap: "Small sample adjusted Akaike's Information Criteria (AICc) model weights and performance scores for linear mixed models predicting total antimicrobial gene burden in White Storks (Ciconia ciconia) in Madrid, Spain.
+"
+#| echo: FALSE
 ModelAIC2 = readRDS(here("4 Results", "Tables", "ModelAIC2.rds"))
 knitr::kable(ModelAIC2)
+
+#| label: Table4
+#| tbl-cap: "Parameter estimates for fixed and random effects parameters and standard error of the highest performing model predicting total antimicrobial gene burden in White Storks in Madrid, Spain."
+#| echo: FALSE
+GLMERmodel1 = readRDS(here("4 Results", "Tables", "LMERmodel2.rds"))
+knitr::kable(GLMERmodel1)
+
 ```

### 2023-04-10 5 Manuscript/manuscript/Manuscript.qmd
Troubleshoot doc

@@
 Binomial generalized linear mixed models (GLMMs) were used to predict multi-drug resistance (MDR) in White Storks, with nest identification classified as a random effects parameter.
-Of these, the highest performing model contained the predictors landfill use index (LUI) and age (@Table1).
-In this model, an increase in the age of the bird was associated with a higher likelihood of MDR being present (@Table2), while an increase in LUI was found to be associated with a decreased likelihood of MDR presence.
+Of these, the highest performing model contained the predictors landfill use index (LUI) and age (Table1).
+In this model, an increase in the age of the bird was associated with a higher likelihood of MDR being present (Table2), while an increase in LUI was found to be associated with a decreased likelihood of MDR presence.
 A second competitive model, with a 97.20% performance score, contained only age as the predictor.
@@
 Linear mixed models (LMMs) were used to predict total antimicrobial gene burden in White Storks, with nest identification again classified as a random effects parameter.
-The global model, containing the variables landfill use index, age, sample period, and nest success, was the highest performing model (@Table3).
-In this model, each variable was positively correlated with antimicrobial gene burden (@Table4).
+The global model, containing the variables landfill use index, age, sample period, and nest success, was the highest performing model (Table3).
+In this model, each variable was positively correlated with antimicrobial gene burden (Table4).
 The next highest performing model contained age as a single predictor, with a model performance score of 86.39%.
@@
 #| echo: FALSE
-GLMERmodel1 = readRDS(here("4 Results", "Tables", "GLMERmodel1.rds"))
+GLMERmodel1 = readRDS(here("4 Results", "Models", "GLMERmodel1.rds"))
 knitr::kable(GLMERmodel1)
@@
 #| echo: FALSE
-GLMERmodel1 = readRDS(here("4 Results", "Tables", "LMERmodel2.rds"))
+GLMERmodel1 = readRDS(here("4 Results", "Models", "LMERmodel2.rds"))
 knitr::kable(GLMERmodel1)

### 2023-04-24 3 Analysis Scripts/Data_Exploratory_Analysis.qmd
Images, Manuscript Edits, etc
Take 1

@@
 cor(variables, use = "complete.obs") #ignore NAs
-corrplot(cor(variables, use = "complete.obs"))
+CorrPlot <- corrplot(cor(variables, use = "complete.obs"))
+CorrPlot
 
@@
 ```{r}
-stork_AMR_clean %>%
+pSB <- stork_AMR_clean %>%
   ggplot(aes(as.factor(yr), sum))+
-  geom_boxplot()
+  geom_boxplot() +
+  ylab('Breeding Season Sampling Period') +
+  xlab ('Antimicrobial Gene Resistance Burden')
+
+while (!is.null(dev.list()))  dev.off()
+#Tiff 
+tiff(file = here('5 Manuscript','manuscript','Images', 'Prelim_BurdenvsSamp.tiff'))
+pSB
+dev.off() 
 ```
@@
 ```{r, message=FALSE, warning=FALSE}
-stork_AMR_2021%>%
+pLN <- stork_AMR_2021%>%
   ggplot(aes(mean_lui, nsuccess, color=as.factor(MDR)))+
   geom_point()+
-  geom_smooth(method = "lm")
+  geom_smooth(method = "lm")+
+  ylab('Nest Success (Includes 2020 to 2021)') +
+  xlab ('Mean Landfill Use Index (LUI)')
+
+#Tiff 
+tiff(file = here('5 Manuscript','manuscript', 'Images', 'Prelim_LUIvsNS.tiff'))
+pLN
+dev.off() 
 ```
@@
 ```{r, warning=FALSE}
-stork_AMR_2021 %>%
+pLM <- stork_AMR_2021 %>%
   ggplot(aes(as.factor(MDR), mean_lui))+
-  geom_boxplot()
+  geom_boxplot()+
+  ylab('Mean Landfill Use Index (LUI)') +
+  xlab ('Multi-Drug Resistance')
+
+#save as tiff
+tiff(file = here('5 Manuscript','manuscript', 'Images', 'Prelim_LUIvsMD.tiff'))
+pLM
+dev.off() 
 ```

### 2023-04-24 3 Analysis Scripts/StatiscalAnalysis_ModelFitting.qmd
readme.md Files + Background
Missing bibTEX files (Currently troubleshooting)
Images, Manuscript Edits, etc
Take 1

@@
 
+##Not working... fix later
+  while (!is.null(dev.list()))  dev.off()
+#Tiff 
+tiff(file = here('5 Manuscript','manuscript','Images', 'LUIvsMDR_LME.tiff'))
 MLp
-
-ggsave(MLp, file="LUIvsMDR_LME.tiff")
+dev.off() 
 ```
@@
 model2 %>% 
-  saveRDS(here("4 Results","Models", "LMERmodel2.rds"))
+  saveRDS(here("4 Results/Models/LMERmodel2.rds"))
 
@@
 
+while (!is.null(dev.list()))  dev.off()
+#Tiff 
+tiff(file = here('5 Manuscript','manuscript','Images', 'LUIvsBurden_LME.tiff'))
 BLp
-
-ggsave(BLp, file="LUIvsBurden_LME.tiff")
+dev.off() 
 ```

### 2023-04-24 5 Manuscript/manuscript/Manuscript.qmd
Images, Manuscript Edits, etc
Take 1

@@
 library(knitcitations)
+library(tidyverse)
 ```
@@
 
-## General Background Information (WILL FIGHT WITH BIBTEX TOMORROW)
+## General Background Information
 
@@
 Resistance genes are thought to alter the fitness of a pathogen [@Friedman2016], and have been correlated with increased virulence [@Escudeiro2019], in turn increasing morbidity and mortality in affected hosts.
-Finally, at the population management level, birds have also been held ‘responsible’ for the dissemination of enteric pathogens (e.g., *Escherichia coli*, *Salmonella* spp.) that have caused outbreaks in produce, yet, they are only moving these bacteria from areas already contaminated by human activities [@Hoelzer2011].
-Thus, understanding how avian behaviors that overlap with risky environmental factors (e.g., foraging in agricultural areas or landfills) impact the carriage of resistance, is key because: 1) it allows us to better understand the role birds play in AMR dissemination, and  2) it informs targeted population management and conservation strategies—such as discouraging birds from utilizing certain areas.
+Finally, at the population management level, birds have also been held 'responsible' for the dissemination of enteric pathogens (e.g., *Escherichia coli*, *Salmonella* spp.) that have caused outbreaks in produce, yet, they are only moving these bacteria from areas already contaminated by human activities [@Hoelzer2011].
+Thus, understanding how avian behaviors that overlap with risky environmental factors (e.g., foraging in agricultural areas or landfills) impact the carriage of resistance, is key because: 1) it allows us to better understand the role birds play in AMR dissemination, and 2) it informs targeted population management and conservation strategies---such as discouraging birds from utilizing certain areas.
 
@@
 It is believed that animals that rely on anthropogenic resources are often concentrated at high densities that promote the large-scale mixing of bacteria, encouranging horizontal gene transfer (HGT) facilitating the rise of novel resistant strains [@Wellington2013].
-For instance, birds that encounter human waste (e.g., by drinking contaminated water) can acquire resistance genes through the exchange of naturally present genes in the bacteria they carry, with the genes in the waste—this can co-select for mobile genetic elements carrying multiple resistant genes [@Wellington2013].
+For instance, birds that encounter human waste (e.g., by drinking contaminated water) can acquire resistance genes through the exchange of naturally present genes in the bacteria they carry, with the genes in the waste---this can co-select for mobile genetic elements carrying multiple resistant genes [@Wellington2013].
 Moreover, ecological factors such as migration and high densities during breeding season may increase acquisition rates and the dissemination of AMR .
@@
 
-## Questions/Hypotheses to be addressed
+## Objective
 
-Does landfill use by storks increase the likelihood of carrying Antimicrobial resistance (AMR) genes and resistance gene burden?.
+Evaluate both antimicrobial gene resistence diversity and burden in white storks that utilize landfills and natural areas over the breeding season of years 2020 to 2021.
 
-Stork that visit the landfill more often will have a higher likelihood of carrying AMR genes and have a higher burden of AMR genes.
+## Hypotheses
 
-Is nest success related to AMR gene burden and multi-drug resistance (i.e., resistance to 3 or more drug classes)?
+Increased landfill use will increase both the diversity and burden of antimicrobial resistance genes in white stork.
 
-Stork with more AMR gene burdens and multi-drug resistance will have lower nest success.
+Nest success will be negatively affected by antimicrobial gene burden and multi-drug resistance (i.e., resistance to 3 or more drug classes).
 
@@
 
-## Data import and cleaning (Reference file and give details HERE. Reference a ReadME/doc here)
+## Data import and cleaning
 
-*Write code that reads in the file and cleans it so it's ready for analysis.
-Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts.
-If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g.
-as supplement in a paper).
-All materials, including files that contain code, should be commented well so everyone can follow along.*
-
-```{r, warning=FALSE}
-library(tidyverse)
-library(here)
-
-```
+Raw data was collected during both field seasons (2020 and 2021) and maintained on an excel file.
+This excel file was amended in 2021 to include the laboratory results from the AMR qPCR.
+The raw excel file can be found in the "1 Data Cleaning Script" folder in this [projects repository](https://github.com/rfranci3/SETHLATTERRAQUELFRANCISCO-MADA-project).
+Cleaned data was evaluated for normality and each variable was standardized and checked correlative relationships.
+All variables were kept as they did not appear strongly correlated.
 
@@
 
-#### (Note: still trying to figure out how to do this with the performance package) We evaluated model fit for all models using Akaike Information Criterion scores adjusted for small sample sizes (AICc: Burnham and Anderson, 2002).
-The best fit model was selected by having a ΔAICc \< 2.\_
-
 {{< pagebreak >}}
@@
 
-## Exploratory/Descriptive analysis (SUMMARY STATS IN EXPLORATION FILE TBD -\> Burden/MDR associated with each class. summary table? with actual summary stats)
+## Exploratory/Descriptive analysis
 
-*Use a combination of text/tables/figures to explore and describe your data.
-Show the most important descriptive results here.
-Additional ones should go in the supplement.
-Even more can be in the R and Quarto files that are part of your project.*
+Cleaned data was then visually explored to evaluate trends.
+Notably, it did not seem to appear that multi-drug resistance played a factor in nest success (Figure 2).
+However, sampling periods did appear important to the amount of ARG burden found in white stork feces (Figure 3).
+The most notable trend in the data however, was the appearance that multi-drug resistence in fact declined with the increase of landfill use by the adult white storks (Figure 4).
+
+![Figure 2: Multi-drug resistance and nest success (i.e., number of chicks fledged) for each nest during the 2020 and 2021 white stork breeding seasons.](Images/Prelim_LUIvsNS.tiff)
+
+![Figure 3: Antimicrobial Drug Resistence burden across all three sampling periods during the 2020 and 2021 white stork breeding seasons.](Images/Prelim_BurdenvsSamp.tiff)
+
+![Figure 4: Multi-drug resistance and Landfill Use Index for each nest during the 2020 and 2021 white stork breeding seasons.](Images/Prelim_LUIvsMD.tiff)
 
@@
 
-```{r}
-#| label: fig-result
-#| fig-cap: "Linear Mixed Model Output with 95% CI."
-#| echo: FALSE
-knitr::include_graphics(here("3 Analysis Scripts","LUIvsBurden_LME.tiff"))
-
-#| label: Figure2
-#| fig-cap: "Multi-drug resistance as predicted by the variables landfill use index and age."
-#| echo: FALSE
-knitr::include_graphics(here("3 Analysis Scripts","LUIvsMDR_LME.tiff"))
-```
-
 ## Full analysis
-
-*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here.*
 
@@
 
-Linear mixed models (LMMs) were used to predict total antimicrobial gene burden in White Storks, with nest identification again classified as a random effects parameter.
-The global model, containing the variables landfill use index, age, sample period, and nest success, was the highest performing model (Table3).
-In this model, each variable was positively correlated with antimicrobial gene burden (Table4).
-The next highest performing model contained age as a single predictor, with a model performance score of 86.39%.
+![Figure 5: Multi-drug resistance as predicted by the variables landfill use index and age during the 2020 and 2021 white stork breeding seasons. The gray is a 95% CI](Images/LUIvsMDR_LME.tiff)
 
@@
 ```
+
+Linear mixed models (LMMs) were used to predict total antimicrobial gene burden in White Storks, with nest identification again classified as a random effects parameter.
+The global model, containing the variables landfill use index, age, sample period, and nest success, was the highest performing model (Table3).
+In this model, each variable was positively correlated with antimicrobial gene burden (Table4).
+The next highest performing model contained age as a single predictor, with a model performance score of 86.39%.
+
+![Figure 6: Anti-microbial resistance gene burden of each nest as predicted by the variables landfill use index and sampling period during the 2020 and 2021 white stork breeding seasons.
+The gray is a 95% CI](Images/LUIvsBurden_LME.tiff)
 
@@
 
-*What are the main take-home messages?*
-
 {{< pagebreak >}}

### 2023-04-25 3 Analysis Scripts/crossvalidation.qmd
Manuscript, supplementary material, start cross-validationn

---
title: "Cross Validation"
format: html
editor: visual
---

# Ciconia ciconia analysis script

#### this script loads the processed, cleaned data, does a simple analysis and saves the results to the results folder

```{r, warning=FALSE, message=FALSE}
#load needed packages. make sure they are installed.
library(ggplot2) #for plotting
library(ggeffects) #to explore plots  
library(broom) #for cleaning up output from lm()
library(broom.mixed) #for cleaning up output from lmer()
library(here) #for data loading/saving
library(multilevelmod) #mixed effect models
library(tidymodels)
library(lme4)
library(sjPlot) #to help plot models
library(performance) #evaluate model fit and performance
library(see)
library(cAIC4)
library(rpart)
library(rpart.plot)
library(glmertree)

#path to data
data_location <- here::here("2 Clean Data","stork_AMR_2021.rds")

#load data. 
data <- readRDS(data_location)
```

```{r}
# Set seed
set.seed(2023)

# Put 3/4 of the data into the training set 
data_split <- initial_split(data, prop = 3/4)

# Create data frames for the two sets
train_data <- training(data_split)
test_data  <- testing(data_split)
```

```{r}
# 5-fold cross-validation, 5 times repeated, using MDR as strata 
folds <- vfold_cv(train_data, v = 5, repeats = 5, strata = as.factor(MDR))

#create model and set engine
tree_model <- 
  decision_tree(
    cost_complexity = tune(),
    tree_depth = tune()) %>% 
  set_engine("rpart") %>% 
  set_mode("regression")

#create recipes for each model

null_mod <-

#create tree workflow
tree_workflow <- workflow() %>%
  add_model(tree_model) %>%
  add_formula(as.factor(MDR) ~ .)
```

```{r}
glmertree_global <- glmertree::glmertree(as.factor(MDR) ~ 1|nes| + s.lui + s.age + samp + s.nsuccess, 
                                         data = train_data, family = "binomial")
plot(glmertree_global)
```

### 2023-04-25 5 Manuscript/manuscript/Manuscript.qmd
Manuscript, supplementary material, start cross-validationn

@@
 
+A total of 126 fecal samples were collected from white storks from 2020 to 2021.
+Of note, resistance genes to tetracyclines and betalactamines were identified in 50 (39.68%) and 73 (57.94%) fecal samples, respectively.
+Multi-drug resistance was identified in 39 (30.95%) samples.
+Mean landfill use index was 0.14 across all samples from 2021, and mean nest success was 55.95% across both sampling periods.
+
 Cleaned data was then visually explored to evaluate trends.
@@
 
-Binomial generalized linear mixed models (GLMMs) were used to predict multi-drug resistance (MDR) in White Storks, with nest identification classified as a random effects parameter.
+Binomial generalized linear mixed models (GLMMs) were used to predict multi-drug resistance (MDR) in white storks, with nest identification classified as a random effects parameter.
 Of these, the highest performing model contained the predictors landfill use index (LUI) and age (Table1).
@@
 
-Linear mixed models (LMMs) were used to predict total antimicrobial gene burden in White Storks, with nest identification again classified as a random effects parameter.
+Linear mixed models (LMMs) were used to predict total antimicrobial gene burden in white storks, with nest identification again classified as a random effects parameter.
 The global model, containing the variables landfill use index, age, sample period, and nest success, was the highest performing model (Table3).
-In this model, each variable was positively correlated with antimicrobial gene burden (Table4).
+In this model, each predictor variable was positively correlated with antimicrobial gene burden (Table4).
 The next highest performing model contained age as a single predictor, with a model performance score of 86.39%.

### 2023-04-25 5 Manuscript/manuscript/supplement/Supplementary_Material.qmd
Manuscript, supplementary material, start cross-validationn

---
title: "Supplementary Material"
subtitle: "A Match Made in Landfills? Exploring the diversity and burden of antimicrobial resistance genes carried by white stork (Ciconia ciconia) throughout the breeding season in Madrid, Spain"
author: Raquel Francisco and Seth Lattner
date: "`r Sys.Date()`"
format:
  pdf:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: EW_AMR.bib
csl: ../../vancouver.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

This shows some materials that could go into a supplementary file. Often you want/need references here too. You can use the same reference bib file for this and the main text (as done here) or have separate bib files.

For illustrative purposes, I'm doing the supplement as pdf. For this to work, you need a (La)TeX system installed. It's easy. Just follow [these steps](https://quarto.org/docs/output-formats/pdf-basics.html).

Of course you would choose the format based on needs.

I'm also using a different style for the references here. (vancouver vs apa in the main manuscript). Usually one would have the formatting of the references the same in those two documents, but I want to illustrate how easy it is to switch reference formatting styles, you just need to get the right CSL file and specify it in the YAML header. We could also have a seperate reference bibtext (`.bib`) file, but here we are using the same.

# Overview

Contained in this document is supplementary material for **"A Match Made in Landfills? Exploring the diversity and burden of antimicrobial resistance genes carried by white stork (*Ciconia ciconia*) throughout the breeding season in Madrid, Spain"**. Supplementary materials include instructions for reproducing data analysis, model results, and any additional analysis.

# Code and file information

To reproduce the data analysis for this manuscript, raw data should first be cleaned by running the DataCleaningScript.qmd file in the "1 Data Cleaning Script" folder. Next, the Data_Exploratory_Analysis.qmd file from the "3 Analysis Scripts" should be run, followed by the StatisticalAnalysis_ModelFitting.qmd file from the same folder. To recreate the manuscript, the Manuscript.qmd file under the "5 Manuscript/manuscript" folders should be run. Raw data is housed in the file StorkAMRraw_2020_2021.csv within the "1 Data Cleaning Script/raw_data" folders. Processed data is housed in the stork_AMR_clean.rds file within the "2 Clean Data" folder. Additionally, a README.md file that serves as a codebook for this project can be found within the "1 Data Cleaning Script/raw_data" folders.

{{< pagebreak >}}

# Additional Method Details

Binomial-family generalized linear mixed models (GLMMs) and linear mixed models (LMMs) were used to assess the effects of multiple predictors on the presence of multi-drug resistance (MDR) and total antimicrobial resistance gene burden (AMR), respectively.

{{< pagebreak >}}

# Additional results

Show additional results here. Those can be some useful exploratory/descriptive figures or tables, or results from additional analyses that didn't make it into the main text.

## Example additional result

@tbl-resulttable1 shows an additional table summarizing a model fit.

```{r}
#| label: tbl-resulttable1
#| tbl-cap: "Another fit table."
#| echo: FALSE
resulttable1 = readRDS(here("results","resulttable1.rds"))
knitr::kable(resulttable1)
```

@fig-result2 shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result2
#| fig-cap: "Height and weight."
#| echo: FALSE
knitr::include_graphics(here("results","height_weight.png"))
```

{{< pagebreak >}}

# Discussion

Any additional discussion regarding the supplementary material/findings.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

{{< pagebreak >}}

# References

### 2023-04-26 3 Analysis Scripts/StatiscalAnalysis_ModelFitting.qmd
continue work on manuscript, supplement, cv

@@
 
-data %>% 
+facet_plot <- data %>% 
   ggplot(aes(x = samp, y = Burden)) + 
@@
   facet_wrap(~ nes) 
+facet_plot
+
+png(file = here('5 Manuscript','manuscript','supplement', 'Supplementary_Material_files', 'facet_plot.png'))
+facet_plot
+dev.off() 
 ```
@@
 
-plot(compare_performance(glmer_fit_null,glmer_fit_global, glmer_fit_1, glmer_fit_2, glmer_fit_3, glmer_fit_4, glmer_fit_5, glmer_fit_6, glmer_fit_7, glmer_fit_8, glmer_fit_9, glmer_fit_10, glmer_fit_11, glmer_fit_12))
+glmer_indices <- plot(compare_performance(glmer_fit_null,glmer_fit_global, glmer_fit_1, glmer_fit_2, glmer_fit_3, glmer_fit_4, glmer_fit_5, glmer_fit_6, glmer_fit_7, glmer_fit_8, glmer_fit_9, glmer_fit_10, glmer_fit_11, glmer_fit_12))
+
+#Tiff 
+png(file = here('5 Manuscript','manuscript','supplement', 'Supplementary_Material_files', 'glmer_indices.png'))
+glmer_indices
+dev.off() 
 ```
@@
 ```{r}
-check_model(glmer_fit_11)
+glmer_check <- check_model(glmer_fit_11)
+
+png(file = here('5 Manuscript','manuscript','supplement', 'Supplementary_Material_files', 'glmer_check.png'))
+glmer_check
+dev.off()
 
@@
            axis.title = c("Landfill Use Index",'Multi-Drug Resistence'))
-
 ##Not working... fix later
@@
 
-plot(compare_performance(lmer_fit2_null, lmer_fit2_global, lmer_fit2_1, lmer_fit2_2, lmer_fit2_3, lmer_fit2_4, lmer_fit2_5, lmer_fit2_6, lmer_fit2_7, lmer_fit2_8,
-                                  lmer_fit2_9, lmer_fit2_10, lmer_fit2_11, lmer_fit2_12))
+lmer_indices <-plot(compare_performance(lmer_fit2_null, lmer_fit2_global, lmer_fit2_1, lmer_fit2_2, lmer_fit2_3, lmer_fit2_4, lmer_fit2_5, lmer_fit2_6, lmer_fit2_7, lmer_fit2_8, lmer_fit2_9, lmer_fit2_10, lmer_fit2_11, lmer_fit2_12))
+
+plot(compare_performance(lmer_fit2_null, lmer_fit2_global, lmer_fit2_1, lmer_fit2_2, lmer_fit2_3, lmer_fit2_4, lmer_fit2_5, lmer_fit2_6, lmer_fit2_7, lmer_fit2_8, lmer_fit2_9, lmer_fit2_10, lmer_fit2_11, lmer_fit2_12))
+
+png(file = here('5 Manuscript','manuscript','supplement', 'Supplementary_Material_files', 'lmer_indices.png'))
+lmer_indices
+dev.off()
+
 ```
@@
 ```{r}
-check_model(lmer_fit2_5)
+lmer_check <- check_model(glmer_fit_global)
+lmer_check
+
+png(file = here('5 Manuscript','manuscript','supplement', 'Supplementary_Material_files', 'lmer_check.png'))
+lmer_check
+dev.off()
 

### 2023-04-26 3 Analysis Scripts/crossvalidation.qmd
continue work on manuscript, supplement, cv

@@
 #path to data
-data_location <- here::here("2 Clean Data","stork_AMR_2021.rds")
+data_location <- here::here("2 Clean Data","stork_AMR_clean.rds")
 
@@
 # 5-fold cross-validation, 5 times repeated, using MDR as strata 
-folds <- vfold_cv(train_data, v = 5, repeats = 5, strata = as.factor(MDR))
+folds <- vfold_cv(train_data, v = 5, repeats = 5, strata = MDR)
 
@@
 
-#create recipes for each model
+#create tree workflow for global model
+tree_workflow_global <- workflow() %>%
+  add_model(tree_model) %>%
+  add_formula(Burden ~ s.lui + s.age + samp + s.nsuccess + nes)
 
-null_mod <-
+#tree grid
+tree_grid <- grid_regular(cost_complexity(),
+                          tree_depth(),
+                          levels = 5)
 
-#create tree workflow
-tree_workflow <- workflow() %>%
+#tree_resamples
+tree_resamp <- 
+  tree_workflow_global %>% 
+  tune_grid(
+    resamples = folds,
+    grid = tree_grid
+    )
+
+tree_resamp %>%
+  collect_metrics()
+
+tree_resamp %>%
+  autoplot()
+```
+```{r}
+#create tree workflow for null
+tree_workflow_null <- workflow() %>%
   add_model(tree_model) %>%
-  add_formula(as.factor(MDR) ~ .)
+  add_formula(Burden ~ 1)
+
+#tree_resamples
+tree_resamp_null <- 
+  tree_workflow_null %>% 
+  tune_grid(
+    resamples = folds,
+    grid = tree_grid
+    )
+
+tree_resamp_null %>%
+  collect_metrics()
+
+tree_resamp_null %>%
+  autoplot()
 ```
-
-```{r}
-glmertree_global <- glmertree::glmertree(as.factor(MDR) ~ 1|nes| + s.lui + s.age + samp + s.nsuccess, 
-                                         data = train_data, family = "binomial")
-plot(glmertree_global)
-```

### 2023-04-26 5 Manuscript/manuscript/supplement/Supplementary_Material.qmd
continue work on manuscript, supplement, cv

@@
 date: "`r Sys.Date()`"
-format:
-  pdf:
-    toc: false
-    number-sections: true
-    highlight-style: github
-bibliography: EW_AMR.bib
-csl: ../../vancouver.csl
+format: html
+editor: visual
 ---
 
-```{r, echo=FALSE, message=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 # load a few R packages
@@
 
-Contained in this document is supplementary material for **"A Match Made in Landfills?
+Contained in this document is supplementary material for the manuscript, **"A Match Made in Landfills?
 Exploring the diversity and burden of antimicrobial resistance genes carried by white stork (*Ciconia ciconia*) throughout the breeding season in Madrid, Spain"**.
-Supplementary materials include instructions for reproducing data analysis, model results, and any additional analysis.
+Supplementary materials include instructions for reproducing data analysis and additonal model results.
 
@@
 
-Show additional results here. Those can be some useful exploratory/descriptive figures or tables, or results from additional analyses that didn't make it into the main text.
+## Exploratory Analysis Results
 
-## Example additional result
+## Full Analysis Results
 
-@tbl-resulttable1 shows an additional table summarizing a model fit.
+A total of 14 GLMMs (including a global and null model) were run to assess the effects of several predictors on MDR.
+Based on Akaike's Information Criterion adjusted for small sample size (AIC*c*), model 11 (containing the predictors LUI and age) was the highest performing model.
+A total of 15 LMMs were run to assess the effects of several predictors on AMR.
+Based on AIC*c*, the global model was the highest performing model, followed by a univariate LMM containing age as the only predictor.
 
-```{r}
-#| label: tbl-resulttable1
-#| tbl-cap: "Another fit table."
-#| echo: FALSE
-resulttable1 = readRDS(here("results","resulttable1.rds"))
-knitr::kable(resulttable1)
-```
+![Figure 1. Facet Plot of single factor repeated measures evaluation the effect of LUI on blaTEM (betalactam) resistence gene burden](Supplementary_Material_files/facet_plot.png)
 
-@fig-result2 shows a scatterplot figure produced by one of the R scripts.
+### GLMM Model Results
 
-```{r}
-#| label: fig-result2
-#| fig-cap: "Height and weight."
-#| echo: FALSE
-knitr::include_graphics(here("results","height_weight.png"))
-```
+![Figure 2. Comparison of Generalized Linear Mixed Model Indices](Supplementary_Material_files/glmer_indices.png)
+
+![Figure 3. Model Check Metrics for Best-performing GLMM](Supplementary_Material_files/glmer_check.png)
+
+### LMM Model Results
+
+![Figure 4. Comparison of Linear Mixed Model Indices](Supplementary_Material_files/lmer_indices.png)
+![Figure 5. Model Check Metrics for Best-performing LMM](Supplementary_Material_files/lmer_check.png)
 

### 2023-04-26 5 Manuscript/manuscript/supplement/Supplementary_Material.tex
continue work on manuscript, supplement, cv

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Supplementary Material},
  pdfauthor={Raquel Francisco and Seth Lattner},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Supplementary Material}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{A Match Made in Landfills? Exploring the diversity and burden
of antimicrobial resistance genes carried by white stork (Ciconia
ciconia) throughout the breeding season in Madrid, Spain}
\author{Raquel Francisco and Seth Lattner}
\date{4/26/23}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners, boxrule=0pt, breakable]}{\end{tcolorbox}}\fi

\begin{verbatim}
Warning: package 'here' was built under R version 4.2.2
\end{verbatim}

\begin{verbatim}
Warning: package 'knitr' was built under R version 4.2.2
\end{verbatim}

This shows some materials that could go into a supplementary file. Often
you want/need references here too. You can use the same reference bib
file for this and the main text (as done here) or have separate bib
files.

For illustrative purposes, I'm doing the supplement as pdf. For this to
work, you need a (La)TeX system installed. It's easy. Just follow
\href{https://quarto.org/docs/output-formats/pdf-basics.html}{these
steps}.

Of course you would choose the format based on needs.

I'm also using a different style for the references here. (vancouver vs
apa in the main manuscript). Usually one would have the formatting of
the references the same in those two documents, but I want to illustrate
how easy it is to switch reference formatting styles, you just need to
get the right CSL file and specify it in the YAML header. We could also
have a seperate reference bibtext (\texttt{.bib}) file, but here we are
using the same.

\hypertarget{overview}{%
\section{Overview}\label{overview}}

Contained in this document is supplementary material for \textbf{``A
Match Made in Landfills? Exploring the diversity and burden of
antimicrobial resistance genes carried by white stork (\emph{Ciconia
ciconia}) throughout the breeding season in Madrid, Spain''}.
Supplementary materials include instructions for reproducing data
analysis, model results, and any additional analysis.

\hypertarget{code-and-file-information}{%
\section{Code and file information}\label{code-and-file-information}}

To reproduce the data analysis for this manuscript, raw data should
first be cleaned by running the DataCleaningScript.qmd file in the ``1
Data Cleaning Script'' folder. Next, the Data\_Exploratory\_Analysis.qmd
file from the ``3 Analysis Scripts'' should be run, followed by the
StatisticalAnalysis\_ModelFitting.qmd file from the same folder. To
recreate the manuscript, the Manuscript.qmd file under the ``5
Manuscript/manuscript'' folders should be run. Raw data is housed in the
file StorkAMRraw\_2020\_2021.csv within the ``1 Data Cleaning
Script/raw\_data'' folders. Processed data is housed in the
stork\_AMR\_clean.rds file within the ``2 Clean Data'' folder.
Additionally, a README.md file that serves as a codebook for this
project can be found within the ``1 Data Cleaning Script/raw\_data''
folders.

\newpage{}

\hypertarget{additional-method-details}{%
\section{Additional Method Details}\label{additional-method-details}}

Binomial-family generalized linear mixed models (GLMMs) and linear mixed
models (LMMs) were used to assess the effects of multiple predictors on
the presence of multi-drug resistance (MDR) and total antimicrobial
resistance gene burden (AMR), respectively.

\newpage{}

\hypertarget{additional-results}{%
\section{Additional results}\label{additional-results}}

A total of 14 GLMMs (including a global and null model) were run to
assess the effects of several predictors on MDR. Based on Akaike's
Information Criterion adjusted for small sample size (AIC\emph{c}),
model 11 (containing the predictors LUI and age) was the highest
performing model. A total of 15 LMMs were run to assess the effects of
several predictors on AMR. Based on AIC\emph{c}, the global model was
the highest performing model, followed by a univariate LMM containing
age as the only predictor.

\hypertarget{lmm-model-results}{%
\subsection{LMM Model Results}\label{lmm-model-results}}

\begin{figure}

{\centering \includegraphics{Images/lmer_indices.tiff}

}

\caption{Table 1. Comparison of Linear Mixed Model Indices}

\end{figure}

\newpage{}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Any additional discussion regarding the supplementary material/findings.

These papers {[}@mckay2020; @mckay2020a{]} are good examples of papers
published using a fully reproducible setup similar to the one shown in
this template.

\newpage{}

\hypertarget{references}{%
\section{References}\label{references}}

\end{document}

### 2023-04-28 Reviews/AnnabellaHinesReviews.qmd

---
title: Project Review Template 
author: Annabella Hines
date: "`r file.mtime(knitr::current_input())`"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: A Match Made in Landfills? Exploring the diversity and burden of antimicrobial resistance genes carried by white stork (Ciconia ciconia) throughout the breeding season in Madrid, Spain

Name of project author(s): Raquel Francisco and Seth Lattner

Name of project reviewer: Annabella Hines

# Specific project content evaluation
Evaluate the different parts of the project by filling in the sections below.

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

I really enjoyed your abstract and background sections; they gave great context to the questions you were trying to answer and helped me understand how it fits into the existing research and work in that field. I like the abundance of information and the references and summaries of previous work

### Summary assessment (PICK ONE, DELETE THE OTHERS)

* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

You had very clear and succinct hypotheses that made sense in the context of the data and the objectives you outlined earlier.

### Summary assessment

* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

There is a codebook provided to explain the variables in the data. The data and its source was described well. If it is feasible, I would've liked a summary table of the data in the manuscript just to reference back to.

### Summary assessment

* source and overall structure of data well explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

I thought the exploration of the data was very thorough and the cleaning as well. The decision to not use 2020 data in the analysis was well explained and supported. Some alternatives were discussed and the most meaningful exploratory results were presented in the manuscript.

### Summary assessment

* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

The basic statistical analysis section was included in the manuscript but it still had the template instructions left in. I see you did do some in the StatisticalAnalysis_ModelFitting.qmd so just choosing the most relevant and moving it over to the manuscript is all you have left. The full analysis statistical models were included and well explained, so I think adding the simple ones may just help the flow to show the steps to how you got to what you did next.

### Summary assessment

* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

The summary and interpretation section could be expanded upon and explained a bit better. The information in parentheses may be better understood by explaining it in a following sentence. I thought the figures were pretty good, I'd just make them more aesthetically pleasing and make the size a bit larger so they are more clearly legible in the manuscript. It seems some figures weren't made or inserted yet as there were captions but just empty space about them.

### Summary assessment

* results are presented ok, with room for improvement

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

The conclusions covered the main results well, however I think they could still be fleshed out a bit more. The limitations seemed to have the general idea but also needs detail. The findings were interpreted appropriately based on the figures and results presented.

### Summary assessment

* minor parts wrong, missing or unclear

## Further comments

Overall I'm really enjoying the lay out and detail of your project. I just think the conclusions and results may need to be more thought out to match the in depth nature of the introduction and background.

# Overall project content evaluation

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

All the files make sense and the junk files were removed or had an explanation as to why they were empty. I like the numbering system in the main directory as it makes it the order to view and run the files very clear.

### Summary assessment

* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

The code and project overall are well documented with adequate supporting files and explanatory information. Each step made sense and had a logical flow from one to the other.

### Summary assessment

* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

I had trouble reproducing some things in you statistical analysis files but I'm pretty that was just a problem on my end or not. When I tried to load certain packages like ggplot2 or see I got errors about the version. Other than that, everything was fully reproducible without edits.

### Summary assessment

* fully reproducible without issues

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

Alternatives were discussed for data processing and models with thorough assessment of their validity and performance. The hypotheses were fully addressed by the statistical tests that were ran.

### Summary assessment

* strong level of thoroughness

## Further comments

Overall you both have done a great job, keep up the good work and little fixes to put the final touches on it!

### 2023-05-01 Reviews/YaoLuReviews.qmd

--- 2023-04-28 Reviews/AnnabellaHinesReviews.qmd
@@
----
-title: Project Review Template 
-author: Annabella Hines
-date: "`r file.mtime(knitr::current_input())`"
-format: 
-  html:
-    toc: true
-    toc-depth: 3
-    number-sections: true
----
-
 # Overview
 
-Title of project: A Match Made in Landfills?
-Exploring the diversity and burden of antimicrobial resistance genes carried by white stork (Ciconia ciconia) throughout the breeding season in Madrid, Spain
+Title of project:Exploring the diversity and burden of antimicrobial resistance genes carried by white stork (Ciconia ciconia) throughout the breeding season in Madrid, Spain
 
-Name of project author(s): Raquel Francisco and Seth Lattner
+Name of project author(s):Raquel Francisco and Seth Latter
 
-Name of project reviewer: Annabella Hines
-
-# Specific project content evaluation
-Evaluate the different parts of the project by filling in the sections below.
+Name of project reviewer:Yao Lu
 
@@
 
-I really enjoyed your abstract and background sections; they gave great context to the questions you were trying to answer and helped me understand how it fits into the existing research and work in that field.
-I like the abundance of information and the references and summaries of previous work
+Good context description. Introduction is detailed.
 
@@
 
-You had very clear and succinct hypotheses that made sense in the context of the data and the objectives you outlined earlier.
+Question(s)/hypotheses are placed in a obvious place.
 
@@
 
-There is a codebook provided to explain the variables in the data.
-The data and its source was described well.
-If it is feasible, I would've liked a summary table of the data in the manuscript just to reference back to.
+Codebook is provided.
 
@@
 
-I thought the exploration of the data was very thorough and the cleaning as well.
-The decision to not use 2020 data in the analysis was well explained and supported.
-Some alternatives were discussed and the most meaningful exploratory results were presented in the manuscript.
+Good exploratory.  Missing check , correlation plot and pairwise check are good.
 
@@
 
-The basic statistical analysis section was included in the manuscript but it still had the template instructions left in.
-I see you did do some in the StatisticalAnalysis_ModelFitting.qmd so just choosing the most relevant and moving it over to the manuscript is all you have left.
-The full analysis statistical models were included and well explained, so I think adding the simple ones may just help the flow to show the steps to how you got to what you did next.
+If there is no specific reason why we should limit the number of variables into 2 or 3.You can try multivariable GLM and finally select a set of variables to fit the model.
+That way should produce a very competitive model to your current glmer_fit_11.
+RMSE should be(99% possibility) less than your current model if not equal to.
+AICs and BICs are not sure but it has high possibility of better.
+
+As your check_model(glmer_fit_11) shows, the Normality of Residuals(QQ plot) shows abnormal.
+It looks like this model is polarization.
+Which means in this model.
+glmer_fit_11 will predict the outcome either overestimate much or underestimate much, even it's the best GLM model compare to the other models you choose.
+This is not we want.
+Since you did much great job in other part.
+If you don't have time to go further at this part.
+Mentioning this abnormal to the readers should be helpful when others want to process you product.
+To address this abnormal, the easiest way is to do transformation to dependent variable or predictors.
+Such as boxcox transformation or basic log transformation.
+
+As your MLp shows, I am not sure how Multi-Drug Resistence measure.
+But in that plot, it looks like a dichotomic variable.
+For dichotomic variables, we usually use mixed effect logistic regression other than linear.
+And this is the reason why the QQ plot show like that.
+When we use linear regression to model dichotomic variable, the QQ plot will looks like the one in 'check_model(glmer_fit_11)'.
 
@@
 
-* strong and reasonable analysis
+* defensible but not optimal analysis
 
@@
 
-The summary and interpretation section could be expanded upon and explained a bit better.
-The information in parentheses may be better understood by explaining it in a following sentence.
-I thought the figures were pretty good, I'd just make them more aesthetically pleasing and make the size a bit larger so they are more clearly legible in the manuscript.
-It seems some figures weren't made or inserted yet as there were captions but just empty space about them.
+Miss table title in page13.
+Also I think you can simplify this table by deleting some columns, such as AICc_wt, since it reflect the same thing as Performance_Score.
+Also you can round it to 2 or 3 digits to make it narrow.
+And show other performance, as you did a great job like in your 'Comparison of Model Indices' plot.
+Why not move those evaluation value to here.
+
+The two table are close which make me think if they are one table at the beginning.
+
+in the second table. page13
+
+Negative sign is separated with the number 4.302130.
 
@@
 
-The conclusions covered the main results well, however I think they could still be fleshed out a bit more.
-The limitations seemed to have the general idea but also needs detail.
-The findings were interpreted appropriately based on the figures and results presented.
+'makes sense because in your prior papers you have found increase nest success with increased LUI' page 14.
+
+Not all audience have that paper.
 
@@
 * minor parts wrong, missing or unclear
-
-## Further comments
-
-Overall I'm really enjoying the lay out and detail of your project.
-I just think the conclusions and results may need to be more thought out to match the in depth nature of the introduction and background.
-
-# Overall project content evaluation
 
@@
 
-All the files make sense and the junk files were removed or had an explanation as to why they were empty.
-I like the numbering system in the main directory as it makes it the order to view and run the files very clear.
+Well structured and easy to follow.
 
@@
 
-The code and project overall are well documented with adequate supporting files and explanatory information. Each step made sense and had a logical flow from one to the other.
+good information and easy to understand.
 
@@
 
-I had trouble reproducing some things in you statistical analysis files but I'm pretty that was just a problem on my end or not.
-When I tried to load certain packages like ggplot2 or see I got errors about the version.
-Other than that, everything was fully reproducible without edits.
+All code can be run without issues.
 
@@
 
-Alternatives were discussed for data processing and models with thorough assessment of their validity and performance. The hypotheses were fully addressed by the statistical tests that were ran.
+Some figures are all white in docx in my computer.
+But I find those in the result folder.
+There are many good plots in the StatiscalAnalysis_ModelFitting_R.
+I think some of those can be processed a little bit and showed in the manuscripts.
+Like 'Comparison of Model Indices' and those statistical plots.
 
@@
 
-* strong level of thoroughness
+* strong level of thorougness
 
@@
 
-Overall you both have done a great job, keep up the good work and little fixes to put the final touches on it!
+Good background discussion. Good exploratory analysis, such as coefficients check, pairwise check. 
+Mixed effect logistic regression maybe more competitive than your current model systematically as one of your outcome looks like dichotomous(I am not sure if that variable should be).
+Variable selection by p-value in multivariable mixed effect model may also produce a more competitive model, if there is no specific reason why limit the number of predictors in 2 or 3.

### 2023-05-05 5 Manuscript/manuscript/supplement/Supplementary_Material.qmd
Finish cross validation, work on supplement

@@
 
+Additionally, a cross-validation approach using tree model was implemented.
+Prior to machine learning analysis, the data were split into training (75%) and testing (25%) groups.
+A 5-fold cross-validation, 5 times repeated approach was implemented using MDR as the strata.
+A tuning grid was utilized using 5 levels of tree depth.
+After fitting the tree model to the training data, the best model was selected using the area under the receiver operating characteristic curve (ROC AUC).
+This final model was then fit to the testing data.
+
 {{< pagebreak >}}
@@
 # Additional results
+
+Contained below are additional results from both the exploratory and full analyses.
 
@@
 
+### Tree Model Results
+
+Cost-complexity parameters for various tree depths are shown in Figure 6.
+
+![Figure 6. Cost-complexity Parameter Plot](Supplementary_Material_files/autoplot.png)
+Results from the finalized tree model can be found in Figure 7.
+
+![Figure 7. Model Check Metrics for Best-performing LMM](Supplementary_Material_files/rpartplot2.png)
+
 {{< pagebreak >}}
@@
 
-Any additional discussion regarding the supplementary material/findings.
-
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
+The tree model was ultimately not incorporated into the final manuscript.
+The ability to account for random effects using the GLMMs and LMMs was better suited for the questions being addressed and the data on hand.
 

### 2023-05-05 3 Analysis Scripts/StatiscalAnalysis_ModelFitting.qmd
Yao Edits
Add images, address analysis, & fix tables

@@
 #Tiff 
-png(file = here('5 Manuscript','manuscript','supplement', 'Supplementary_Material_files', 'glmer_indices.png'))
+tiff(file = here('5 Manuscript','manuscript','Images', 'glmer_indices.tiff'))
 glmer_indices
@@
 
-Use AIC to see the best model; This doesnt look great either...
+Use AIC to see the best model
 
@@
 
-png(file = here('5 Manuscript','manuscript','supplement', 'Supplementary_Material_files', 'lmer_indices.png'))
+tiff(file = here('5 Manuscript','manuscript','Images', 'lmer_indices.tiff'))
 lmer_indices

### 2023-05-05 5 Manuscript/manuscript/Manuscript.qmd
Addressing Annabella's comment about figures
Figures now populated on manuscript
Yao Edits
Add images, address analysis, & fix tables

@@
 
-## Exploratory/Descriptive analysis
+## Exploratory preliminary analysis
 
@@
 However, sampling periods did appear important to the amount of ARG burden found in white stork feces (Figure 3).
-The most notable trend in the data however, was the appearance that multi-drug resistence in fact declined with the increase of landfill use by the adult white storks (Figure 4).
+The most notable trend in the data however, was the appearance that multi-drug resistance in fact declined with the increase of landfill use by the adult white storks (Figure 4).
 
@@
 
-## Basic statistical analysis
+## Full Statistical Analysis
 
-*To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
-simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*
-
-## Full analysis
+![Figure 5: Multi-drug resistance as predicted by the variables landfill use index and age during the 2020 and 2021 white stork breeding seasons. The gray is a 95% CI.](Images/LUIvsMDR_LME.tiff)
 
 Binomial generalized linear mixed models (GLMMs) were used to predict multi-drug resistance (MDR) in white storks, with nest identification classified as a random effects parameter.
-Of these, the highest performing model contained the predictors landfill use index (LUI) and age (Table1).
-In this model, an increase in the age of the bird was associated with a higher likelihood of MDR being present (Table2), while an increase in LUI was found to be associated with a decreased likelihood of MDR presence.
-A second competitive model, with a 97.20% performance score, contained only age as the predictor.
+Of these, the highest performing model contained the predictors landfill use index (LUI) and age (Table 1).
+In this model, an increase in the age of the bird was associated with a higher likelihood of MDR being present (Table 2), while an increase in LUI was found to be associated with a decreased likelihood of MDR presence.
+A second competitive model, with a 97.20% performance score, contained only age as the predictor.In the analysis, several models where overfit, not including the top two models.
+However, this is likely due to the small sample size and the restricted range of the standardized variables.
 
-![Figure 5: Multi-drug resistance as predicted by the variables landfill use index and age during the 2020 and 2021 white stork breeding seasons. The gray is a 95% CI](Images/LUIvsMDR_LME.tiff)
+```{r}
+#| label: Table2
+#| tbl-cap: "Parameter estimates for fixed and random effects parameters and standard error of the highest performing model predicting multi-drug resistance in White Storks in Madrid, Spain."
+#| echo: FALSE
+GLMERmodel1 = readRDS(here("4 Results", "Models", "GLMERmodel1.rds"))
+knitr::kable(GLMERmodel1)
+```
+
+![Figure 6: Visual model performance analysis of the all the GLMMs used to evaluate Multi-Drug Resistance (i.e., resistance to 3 or more drug classes).
+Covariates used includes: Age, Sampling Period, Landfill Use, and Nest Success.
+Individual Nests were the fixed variable.
+Due to small sample sizes AICc was viewed as the most appropriate metric, however AIC, BIC, RSME, R2, and Sigma are displayed.](Images/glmer_indices.tiff)
 
@@
 
-#| label: Table2
-#| tbl-cap: "Parameter estimates for fixed and random effects parameters and standard error of the highest performing model predicting multi-drug resistance in White Storks in Madrid, Spain."
-#| echo: FALSE
-GLMERmodel1 = readRDS(here("4 Results", "Models", "GLMERmodel1.rds"))
-knitr::kable(GLMERmodel1)
-
 ```
 
+![Figure 7: Anti-microbial resistance gene burden of each nest as predicted by the variables landfill use index and sampling period during the 2020 and 2021 white stork breeding seasons.
+The gray is a 95% CI.](Images/LUIvsBurden_LME.tiff)
+
 Linear mixed models (LMMs) were used to predict total antimicrobial gene burden in white storks, with nest identification again classified as a random effects parameter.
-The global model, containing the variables landfill use index, age, sample period, and nest success, was the highest performing model (Table3).
-In this model, each predictor variable was positively correlated with antimicrobial gene burden (Table4).
+The global model, containing the variables landfill use index, age, sample period, and nest success, was the highest performing model (Table 3).
+In this model, each predictor variable was positively correlated with antimicrobial gene burden (Table 4).
 The next highest performing model contained age as a single predictor, with a model performance score of 86.39%.
+Similarly to the top analysis, several models where overfit likely due to similar reasons.
 
-![Figure 6: Anti-microbial resistance gene burden of each nest as predicted by the variables landfill use index and sampling period during the 2020 and 2021 white stork breeding seasons.
-The gray is a 95% CI](Images/LUIvsBurden_LME.tiff)
+```{r}
+#| label: Table4
+#| tbl-cap: "Parameter estimates for fixed and random effects parameters and standard error of the highest performing model predicting total antimicrobial gene burden in White Storks in Madrid, Spain."
+#| echo: FALSE
+GLMERmodel1 = readRDS(here("4 Results", "Models", "LMERmodel2.rds"))
+knitr::kable(GLMERmodel1)
+```
+
+![Figure 8: Visual model performance analysis of the all the LMMs used to evaluate Antimicrobial Resistance Gene Burden.
+Covariates used includes: Age, Sampling Period, Landfill Use, and Nest Success.
+Individual Nests were the fixed variable.
+Due to small sample sizes AICc was viewed as the most appropriate metric, however AIC, BIC, RSME, R2, and Sigma are displayed.](Images/lmer_indices.tiff)
 
@@
 
-#| label: Table4
-#| tbl-cap: "Parameter estimates for fixed and random effects parameters and standard error of the highest performing model predicting total antimicrobial gene burden in White Storks in Madrid, Spain."
-#| echo: FALSE
-GLMERmodel1 = readRDS(here("4 Results", "Models", "LMERmodel2.rds"))
-knitr::kable(GLMERmodel1)
-
 ```
@@
 
--   As multi-drug resistance and class diversity increase (similar things I know) throughout the breeding season (compounding effect likely), mean LUI decreases and nest success decreases (makes sense because in your prior papers you have found increase nest success with increased LUI).
+-   As multi-drug resistance increases throughout the breeding season, mean LUI decreases and nest success decreases.
+There is a positive effect between increased LUI and nest success [@Lopez-Garcia2021].
+Intuitively this makes some sense due to MDR being associated with increased anthropogenic exposure, heavy metal exposure, and increased virulence of pathogenic bacteria.
+More target research would be needed to better understand how MDR may effect nest success long-term.
 
--   Resistance gene burden appears to increase as mean LUI use and age increase.
-o Most of the burden is due to blaTEM, a common resistance gene associated with anthropogenic impact.
-Sampling period does not appear to explain burden, but the top blaTEM model did show a trend in burden increasing from the 2nd sampling period to the 3rd sampling period and age.
+-   Resistance gene burden appears to increase as mean LUI decreases and age increases.
+Most of the burden is due to blaTEM, a common resistance gene associated with beta-lactam resistance.
+Beta-lactam resistance is one of the most studied AMRs in wildlife.
+Drugs within the Beta-lactam class are often first line antibiotics, such as penicillin, and are active against many gram-positive, gram-negative, and anaerobic organisms.
+At this point, beta-lactam resistance is thought to be prevalent on most landscapes due to human or livestock encroachment thus is often used as a measure of anthropogenic impact.
+While sampling periods did not explain burden, the top blaTEM model did show a trend in burden increasing from the 2nd sampling period to the 3rd sampling period.This may be related to shear increase of foraging required to feed ~3 to 4 kilogram fledglings.This foraging could have occurred in areas other then the landfill, such as neighboring cattle pastures which may also be sources for ARGs.
 
@@
 
--   Not much statistically significant data, thus we may have to argue biological significance.
+-   While none of our models resulted in statistically significant data, valuable trends can be seen in respect to the relationship between breeding season events, AMR trends, and landfill use.
+Hopefully, these trends can foster future research to consider expandings beyond the hypothesis that landfills are the risky place for wildlife to acquire ARGs.
+
+-   The dataset utilized for the models was much smaller then previously envisioned due to COVID-19 barring landfill data collection throughout the entirety of 2020.
+Perhaps with more observations, the results of the models may have been stronger and variables may have been more normally distributed.
+Moving forward, it may be best to transform our standardized variables using a log transformation to attempt to achieve a more normal distribution for analysis, especially in regard to our GLMMs.
 
@@
 LUI appears to be correlated with higher levels of AMR gene burden in storks.
-As LUI increases thorough the breeding season [@Bialas2020] resistance gene burden also increases with beta lactam resistance contributing to the majority of the burden.
-However, multidrug resistance appears to decrease as LUI increases, thus it is likely that storks are being exposed to antimicrobial resistance genes at other foraging areas (urban centers, agricultural pastures, etc.).
+As LUI increases thorough the breeding season [@Bialas2020] resistance gene burden also increases with beta-lactam resistance contributing to the majority of the burden.
+However, multi-drug resistance appears to decrease as LUI increases, thus it is likely that storks are being exposed to antimicrobial resistance genes at other foraging areas (urban centers, agricultural pastures, etc.).
 Our results suggest that landfills may not be contributing significantly to the emergence and maintenance of AMR in this system.

### 2023-05-05 3 Analysis Scripts/crossvalidation.qmd

@@
   add_model(tree_model) %>%
-  add_formula(Burden ~ s.lui + s.age + samp + s.nsuccess + nes)
+  add_formula(Burden ~ s.lui + s.age + samp + s.nsuccess)
 
@@
   autoplot()
+
+#select best tree model
+best_tree <- tree_resamp %>%
+  select_best(metric = 'rmse')
+
+#final model
+final_tree <- tree_workflow_global %>%
+  finalize_workflow(best_tree)
+
+#fit final model to data
+tree_fit_final <- final_tree %>%
+  fit(train_data)
+
+#plot final fit
+rpart.plot::rpart.plot(extract_fit_parsnip(tree_fit_final)$fit)
+
+#tree residual plot
+tree_pred <- tree_fit_final %>%
+  fit(train_data) %>%
+  predict(train_data)
+
+tree_resid <- train_data$Burden - tree_pred$.pred
+tree_df <- tibble(tree_resid, tree_pred)
+
+ggplot(aes(.pred, tree_resid), data = tree_df)+
+  geom_point()+
+  labs(x = "Predictions", y = "Residuals")+
+  geom_hline(yintercept = 0)+
+  theme_bw()
 ```
+
 ```{r}
-#create tree workflow for null
-tree_workflow_null <- workflow() %>%
-  add_model(tree_model) %>%
-  add_formula(Burden ~ 1)
+#create model and set engine
+tree_model2 <- 
+  decision_tree(
+    cost_complexity = tune(),
+    tree_depth = tune()) %>% 
+  set_engine("rpart") %>% 
+  set_mode("classification")
+
+#create tree workflow for global model
+tree_workflow_global2 <- workflow() %>%
+  add_model(tree_model2) %>%
+  add_formula(as.factor(MDR) ~ s.lui + s.age + samp + s.nsuccess)
 
 #tree_resamples
-tree_resamp_null <- 
-  tree_workflow_null %>% 
+tree_resamp2 <- 
+  tree_workflow_global2 %>% 
   tune_grid(
@@
 
-tree_resamp_null %>%
+tree_resamp2 %>%
   collect_metrics()
 
-tree_resamp_null %>%
+tree_resamp2 %>%
   autoplot()
+
+#select best tree model
+best_tree2 <- tree_resamp2 %>%
+  select_best()
+
+#final model
+final_tree2 <- tree_workflow_global2 %>%
+  finalize_workflow(best_tree2)
+
+#fit final model to data
+tree_fit_final2 <- final_tree2 %>%
+  fit(train_data)
+
+#plot final fit
+rpart.plot::rpart.plot(extract_fit_parsnip(tree_fit_final2)$fit)
+
+#tree residual plot
+tree_pred2 <- tree_fit_final2 %>%
+  fit(train_data) %>%
+  predict(train_data)
+
+tree_resid2 <- train_data$MDR - tree_pred$.pred
+tree_df2 <- tibble(tree_resid2, tree_pred2)
 ```

### 2023-05-05 5 Manuscript/manuscript/Manuscript.qmd
Manuscript edits

@@
 This supports the belief that birds who utilize and move between human-dominated and natural landscapes may become both reservoirs and efficient transporters of resistance determinants to natural environments.For wildlife in general, resistance prevalence tends to increase with proximity to human populations [@Niu2013].
-It is believed that animals that rely on anthropogenic resources are often concentrated at high densities that promote the large-scale mixing of bacteria, encouranging horizontal gene transfer (HGT) facilitating the rise of novel resistant strains [@Wellington2013].
+It is believed that animals that rely on anthropogenic resources are often concentrated at high densities that promote the large-scale mixing of bacteria, encouraging horizontal gene transfer (HGT) facilitating the rise of novel resistant strains [@Wellington2013].
 For instance, birds that encounter human waste (e.g., by drinking contaminated water) can acquire resistance genes through the exchange of naturally present genes in the bacteria they carry, with the genes in the waste---this can co-select for mobile genetic elements carrying multiple resistant genes [@Wellington2013].
@@
 
-Evaluate both antimicrobial gene resistence diversity and burden in white storks that utilize landfills and natural areas over the breeding season of years 2020 to 2021.
+Evaluate both antimicrobial gene resistance diversity and burden in white storks that utilize landfills and natural areas over the breeding season of years 2020 to 2021.
 
@@
 We are evaluating for the presence and burden of antimicrobial resistance genes in white stork feces.
-Codebook is a WIP located in raw data folder.
+A codebook containing descriptions of variables and their abbreviations can be found in the raw data folder, located within the project repository.
 
@@
 In this model, an increase in the age of the bird was associated with a higher likelihood of MDR being present (Table 2), while an increase in LUI was found to be associated with a decreased likelihood of MDR presence.
-A second competitive model, with a 97.20% performance score, contained only age as the predictor.In the analysis, several models where overfit, not including the top two models.
+A second competitive model, with a 97.20% performance score, contained only age as the predictor.
+In this analysis, several models where overfit, not including the top two models.
 However, this is likely due to the small sample size and the restricted range of the standardized variables.
@@
 The next highest performing model contained age as a single predictor, with a model performance score of 86.39%.
-Similarly to the top analysis, several models where overfit likely due to similar reasons.
+As with the GLMM analysis, several models, including the top model, where overfit, likely due to similar reasons.
 
@@
 There is a positive effect between increased LUI and nest success [@Lopez-Garcia2021].
-Intuitively this makes some sense due to MDR being associated with increased anthropogenic exposure, heavy metal exposure, and increased virulence of pathogenic bacteria.
-More target research would be needed to better understand how MDR may effect nest success long-term.
+Intuitively, this makes some sense due to MDR being associated with increased anthropogenic exposure, heavy metal exposure, and increased virulence of pathogenic bacteria.
+More targeted research would be needed to better understand how MDR may effect nest success long-term.
 
 -   Resistance gene burden appears to increase as mean LUI decreases and age increases.
-Most of the burden is due to blaTEM, a common resistance gene associated with beta-lactam resistance.
+Most of the total burden is attributed to blaTEM, a common resistance gene associated with beta-lactam resistance.
 Beta-lactam resistance is one of the most studied AMRs in wildlife.
-Drugs within the Beta-lactam class are often first line antibiotics, such as penicillin, and are active against many gram-positive, gram-negative, and anaerobic organisms.
-At this point, beta-lactam resistance is thought to be prevalent on most landscapes due to human or livestock encroachment thus is often used as a measure of anthropogenic impact.
-While sampling periods did not explain burden, the top blaTEM model did show a trend in burden increasing from the 2nd sampling period to the 3rd sampling period.This may be related to shear increase of foraging required to feed ~3 to 4 kilogram fledglings.This foraging could have occurred in areas other then the landfill, such as neighboring cattle pastures which may also be sources for ARGs.
+Drugs within the Beta-lactam class are often first-line antibiotics, such as penicillin, and are active against many gram-positive, gram-negative, and anaerobic organisms.
+At this point, beta-lactam resistance is thought to be prevalent on most landscapes due to human or livestock encroachment and thus is often used as a measure of anthropogenic impact.
+While sampling periods did not explain burden, the top blaTEM model did show a trend in burden increasing from the 2nd sampling period to the 3rd sampling period.This may be related to sheer increase of foraging required to feed \~3 to 4 kilogram fledglings.This foraging could have occurred in areas other then the landfill, such as neighboring cattle pastures which may also be sources for ARGs.
 
@@
 -   While none of our models resulted in statistically significant data, valuable trends can be seen in respect to the relationship between breeding season events, AMR trends, and landfill use.
-Hopefully, these trends can foster future research to consider expandings beyond the hypothesis that landfills are the risky place for wildlife to acquire ARGs.
+Hopefully, these trends can foster future research to consider expanding beyond the hypothesis that landfills are the risky place for wildlife to acquire ARGs.
 
@@
 Perhaps with more observations, the results of the models may have been stronger and variables may have been more normally distributed.
+While model residuals should be normally-distributed, model performance metrics indicated polarization in the residuals in several of our top model, which can lead to under- and over-fitting data.
 Moving forward, it may be best to transform our standardized variables using a log transformation to attempt to achieve a more normal distribution for analysis, especially in regard to our GLMMs.

### 2023-05-05 5 Manuscript/manuscript/supplement/Supplementary_Material.qmd
Final edits to supplement

@@
 ```
-
-This shows some materials that could go into a supplementary file.
-Often you want/need references here too.
-You can use the same reference bib file for this and the main text (as done here) or have separate bib files.
-
-For illustrative purposes, I'm doing the supplement as pdf.
-For this to work, you need a (La)TeX system installed.
-It's easy.
-Just follow [these steps](https://quarto.org/docs/output-formats/pdf-basics.html).
-
-Of course you would choose the format based on needs.
-
-I'm also using a different style for the references here.
-(vancouver vs apa in the main manuscript).
-Usually one would have the formatting of the references the same in those two documents, but I want to illustrate how easy it is to switch reference formatting styles, you just need to get the right CSL file and specify it in the YAML header.
-We could also have a seperate reference bibtext (`.bib`) file, but here we are using the same.
 
@@
 
-Contained below are additional results from both the exploratory and full analyses.
-
-## Exploratory Analysis Results
+Contained below are additional results from the full analyses.
 
@@
 
-![Figure 4. Comparison of Linear Mixed Model Indices](Supplementary_Material_files/lmer_indices.png)
-![Figure 5. Model Check Metrics for Best-performing LMM](Supplementary_Material_files/lmer_check.png)
+![Figure 4.
+Comparison of Linear Mixed Model Indices](Supplementary_Material_files/lmer_indices.png) ![Figure 5.
+Model Check Metrics for Best-performing LMM](Supplementary_Material_files/lmer_check.png)
 
@@
 
-![Figure 6. Cost-complexity Parameter Plot](Supplementary_Material_files/autoplot.png)
-Results from the finalized tree model can be found in Figure 7.
+![Figure 6. Cost-complexity Parameter Plot](Supplementary_Material_files/autoplot.png) Results from the finalized tree model can be found in Figure 7.
 
